---
title: "LogicComm: Multi-Sample Cell-Cell Communication Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LogicComm: Multi-Sample Cell-Cell Communication Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Overview

**LogicComm** approaches cell-cell communication differently from CellChat or
CellPhoneDB. Instead of requiring pre-annotated cell types and building
communication probability matrices, LogicComm:

1. Converts each cell's expression profile to a **binary logic state** using
   Relative Expression Ordering (REO) — no cross-sample normalization needed.

2. Detects **Logic Consensus Scores (LCS)** per sample: for each L-R pair,
   how often is the ligand logic "ON" in a sender cell AND the receptor logic
   "ON" in a neighboring receiver cell?

3. Compares LCS across **N Case vs N Control** samples using asymmetric
   frequency voting — no cell type matching required across samples.

This makes LogicComm ideal for cohort-level studies where cell type
proportions differ across samples and conventional matching is unreliable.

---

## Installation

```{r install, eval=FALSE}
# Core dependencies
install.packages(c(
  "Matrix", "dplyr", "ggplot2", "ggrepel", "scales", "Rcpp",
  "testthat", "rmarkdown"
))

# Optional dependencies
install.packages("pheatmap")              # heatmap plotting
install.packages(c("Seurat", "SeuratObject"))  # Seurat input / KNN extraction

# Install from local source tarball (adjust the version number)
install.packages("LogicComm_0.10.4.tar.gz", repos = NULL, type = "source")
# Or a published release: github.com/awaragml00029-debug/LogicComm/releases
```

Input matrices should be **genes x cells**. Sparse matrices are supported. By default, REO anchors are computed from all genes in each cell, while `lr_genes` controls only which genes are returned and scored. On Windows, keep `mc_cores = 1` in `run_multisample()` unless you provide another parallel backend.

`lr_pairs_human` is a compact human LR seed database compiled from public resources including CellChatDB, CellPhoneDB, and NicheNet. CellChatDB itself is a manually curated, literature-supported database; in the CellChat repository the full human database is distributed as `data/CellChatDB.human.rda`. LogicComm converts selected interactions into a simpler table with explicit `ligand_genes` and `receptor_genes` list columns for binary REO/LCS scoring.

---

## Simulated Data

We'll simulate 8 samples (4 Case + 4 Control) with ~300 cells each. The Case
group has elevated VEGFA and IL6 signaling.

```{r simulate}
library(LogicComm)
set.seed(2025)

n_cells   <- 300
n_genes   <- 80
gene_pool <- c(
  "VEGFA","KDR","FLT1","TGFB1","TGFBR1","TGFBR2",
  "CXCL12","CXCR4","TNF","TNFRSF1A","IL6","IL6R","IL6ST",
  "COL1A1","CD44","PDGFB","PDGFRB","FGF2","FGFR1",
  "CD274","PDCD1","HGF","MET","ANGPT1","TEK",
  "CCL2","CCR2","CXCL10","CXCR3",
  paste0("BG", seq_len(n_genes - 28))  # background genes
)

make_sample <- function(seed, case = FALSE) {
  set.seed(seed)
  mat <- matrix(stats::rpois(length(gene_pool) * n_cells, lambda = 1.2),
                nrow = length(gene_pool), ncol = n_cells,
                dimnames = list(gene_pool, paste0("C", seed, "_", seq_len(n_cells))))
  if (case) {
    # Elevate VEGF and IL6 axes in Case
    mat["VEGFA", ] <- mat["VEGFA", ] + rpois(n_cells, 6)
    mat["KDR",   ] <- mat["KDR",   ] + rpois(n_cells, 4)
    mat["IL6",   ] <- mat["IL6",   ] + rpois(n_cells, 5)
    mat["IL6R",  ] <- mat["IL6R",  ] + rpois(n_cells, 3)
  }
  mat
}

# 8 samples: 4 Case + 4 Control
sample_list <- c(
  lapply(1:4,  function(s) make_sample(s,     case = TRUE)),
  lapply(5:8,  function(s) make_sample(s + 10, case = FALSE))
)
names(sample_list) <- c(paste0("Case_", 1:4), paste0("Ctrl_", 1:4))

group_info <- c(rep("Case", 4), rep("Ctrl", 4))
names(group_info) <- names(sample_list)
```

---

## Step 1: Build the REO Matrix

For each sample, convert expression to binary logic:

```{r reo}
lr_genes <- all_lr_genes(lr_pairs_human)

# Single sample example
reo_demo <- calc_REO_matrix(
  sample_list[[1]],
  lr_genes       = lr_genes,
  rank_threshold = 0.5   # median anchor
)
reo_demo
```

The REO matrix is sparse: `1` means a gene is expressed above the cell's
median background; `0` means it falls below. Importantly, `lr_genes` does **not**
define the background by default. The anchor is computed from all genes in the
sample matrix, then only retained LR genes are converted into the output REO
matrix and scored. This keeps REO states comparable when you switch LR databases.

**Key insight:** This within-cell comparison makes the logic state
*independent of sequencing depth* — a cell with 2,000 UMIs and one with
10,000 UMIs are judged by the same relative standard.

---

## Step 2: Logic Consensus Score (LCS) per Sample

```{r lcs}
# Dataset-level L-R co-expression score (no graph required)
lcs_demo <- IdentifyLogicConsensus(reo_demo, verbose = FALSE)

# Top 10 active L-R pairs
sort(lcs_demo, decreasing = TRUE)[1:10]
```

> **v0.12 note.** LogicComm scores communication from REO co-expression and no
> longer uses a per-cell KNN/SNN neighborhood graph. Legacy neighborhood
> arguments (`knn_mat`, `mode`, `graph_name`, ...) are accepted but deprecated
> and ignored with a warning. Cell-type-resolved discovery uses
> `discover_celltype_communication()` (see below).

---

## Step 3: Multi-Sample Comparison (One-Stop Pipeline)

```{r multisample, message=TRUE}
result <- run_multisample(
  sample_list,
  group_info     = group_info,
  lr_db          = lr_pairs_human,
  rank_threshold = 0.5,
  lcs_threshold  = 0.01,
  mc_cores       = 1,
  verbose        = TRUE
)
```

```{r print_result}
# Top enriched pairs
print(result$comparison, n = 10)
```

---

## Step 4: Visualize — Bubble Plot

```{r bubble, fig.cap="Case vs Control frequency bubble plot. Points above the diagonal are enriched in Case. Bubble size = |asymmetry|."}
plot_lcs_bubble(result$comparison, top_n = 15, min_asymmetry = 0.2)
```

---

## Step 5: Heatmap Across Samples

```{r heatmap, fig.height=6, eval=requireNamespace("pheatmap", quietly = TRUE), fig.cap="LCS values across all 8 samples for the top differentially active pairs (rendered when the suggested pheatmap package is installed)."}
plot_lcs_heatmap(result$comparison, group_info = group_info,
                 top_n = 20, fdr_cutoff = 1)  # show top demo pairs even when the small simulation is underpowered
```

---

## Step 6: Gene Rank Shift Analysis

Identifies genes that became **transcriptionally dominant** in Case cells —
even if their fold-change is modest.

```{r rank_shift}
rs <- calc_rank_shift(
  sample_list,
  group_info = group_info,
  genes      = lr_genes
)
print(rs, n = 8)
```

```{r rank_shift_plot, fig.cap="Rank-shift volcano plot. X-axis = how much a gene's within-cell dominance changed. Genes to the right became more prominent in Case cells."}
plot_rank_shift(rs, top_n_label = 15, min_shift_label = 0.05)
```

---

## Step 7: Identify Driving Clusters

If you have cell type annotations, identify which clusters are responsible for
the top signaling pairs:

```{r clusters, eval=FALSE}
# Assign mock cluster labels
cluster_labels <- setNames(
  rep(c("Fibroblast","Endothelial","Macrophage"), length.out = n_cells),
  colnames(sample_list[[1]])
)

labels <- AutoLabelLogicClusters(
  reo_demo,
  cell_labels = cluster_labels,
  lr_pairs    = c("VEGFA_KDR", "IL6_IL6R", "TGFB1_TGFBR1")
)
print(labels)
```


---

## Step 8: Cell-Type Communication and Signaling Roles

When cell-type labels are available, LogicComm can summarize LCS by directed
celltypeA -> celltypeB communication. This answers biological questions such as
how many active interactions occur between two cell types, how strong the summed
logic-consensus signal is, whether a finding is supported by enough cell-cell
edges, and whether each cell type behaves more like a sender, receiver, mediator,
or influencer.

```{r celltype_comm_intro, eval=FALSE}
cluster_labels <- setNames(
  rep(c("Fibroblast", "Endothelial", "Macrophage"), length.out = n_cells),
  colnames(sample_list[[1]])
)

ct_comm <- summarize_celltype_communication(
  reo_demo,
  cell_labels   = cluster_labels,   # scored from cell-type co-expression (no graph)
  lr_db         = lr_pairs_human,
  lcs_threshold = 0.01,
  min_edges     = 20
)

# Interaction quantity, strength, and edge support
plot_celltype_heatmap(ct_comm, metric = "n_active_lr")
plot_celltype_heatmap(ct_comm, metric = "sum_lcs")
plot_celltype_heatmap(ct_comm, metric = "sum_active_edges")

# Role positioning, role profile, and mechanistic explanation
plot_celltype_roles(ct_comm)
plot_celltype_role_heatmap(ct_comm)
plot_lr_bubble_by_celltype(ct_comm, sender = "Macrophage", receiver = "Endothelial")
plot_lr_activity_balance(ct_comm, sender = "Macrophage", receiver = "Endothelial")
```

`role_summary` contains four centrality-inspired role scores, disambiguated count
fields, role-separation labels, and communication-evidence labels:

```{r role_summary_intro, eval=FALSE}
ct_comm$role_summary[, c("cell_type", "sender_role_score", "receiver_role_score",
                         "mediator_role_score", "influencer_role_score",
                         "outgoing_lr_event_count", "incoming_lr_event_count",
                         "outgoing_target_type_count", "incoming_source_type_count",
                         "dominant_role", "role_separation_label", "communication_evidence_label", "role_reliability_label")]
```

- `sender_role_score`: outgoing logic-consensus strength.
- `receiver_role_score`: incoming logic-consensus strength.
- `mediator_role_score`: betweenness / gatekeeper tendency.
- `influencer_role_score`: information-flow tendency.
- `dominant_role = "Low-communication"` means the cell type did not pass the
  minimum hub/event filters for confident role assignment.

For the recommended end-to-end discovery (specificity + proliferation filter +
axis-level permutation FDR), use the one-call pipeline:

```{r discover_intro, eval=FALSE}
reo_rk <- calc_REO_matrix(count_matrix, lr_genes = all_lr_genes(lr_pairs_human),
                          return_rank = TRUE)
res <- discover_celltype_communication(
  reo_mat = reo_rk, cell_labels = cluster_labels, lr_db = lr_pairs_human,
  expr = count_matrix, lcs_weighting = "rank",
  n_perm = 1000, n_cores = 8, seed = 1
)
res$shortlist   # axes passing permutation_fdr <= 0.1
```

For cohort analysis, calculate `ct_comm` inside each sample and convert it to a
named vector with `celltype_comm_to_lcs()` before using `CompareLogicGroups()`.
Differential cell-type communication can then be visualized with
`plot_differential_celltype_heatmap()` and `plot_differential_celltype_volcano()`.

---

## Step 9: Per-Cell Communication Score

Score every cell by how many L-R pairs it participates in:

```{r cell_score}
scores <- score_lr_activity(reo_demo, mode = "both", aggregate = TRUE, verbose = FALSE)
summary(scores$comm_score)
```

```{r cell_score_hist}
hist(scores$comm_score, breaks = 30, col = "#64B5F6",
     main = "Per-Cell Communication Score Distribution",
     xlab = "Fraction of L-R pairs active", border = "white")
```

Top communicating cells:

```{r rank_cells}
top_cells <- rank_comm_cells(scores, n = 10)
print(top_cells[, c("cell","sender_score","receiver_score","comm_score","rank")])
```

`rank_comm_cells()` requires a combined `comm_score`; use `score_lr_activity(..., mode = "both", aggregate = TRUE)` before ranking cells.

---


---

## Step 10: Publication-Level Extensions

For full Seurat/PBMC examples, see `Seurat_demo.Rmd`. The same simulated objects
can still demonstrate the new publication-oriented helpers.

### Cell-type QC and markdown report

```{r publication_qc, eval=FALSE}
findings <- summarize_communication_findings(ct_comm, top_n = 10)
findings

plot_communication_qc(ct_comm)
plot_role_confidence(ct_comm)

write_communication_report(ct_comm, file = "LogicComm_demo_report.md")
```

### Sample-level GLM for cell-type communication

When multiple biological samples have cell-type communication summaries, model
active edge counts with sample metadata rather than treating cells as replicates.

```{r sample_level_glm, eval=FALSE}
# sample_ct_list: named list of LogicCommCellTypeComm objects.
# sample_metadata: data frame with rownames matching sample names.
glm_res <- fit_celltype_comm_glm(
  sample_ct_list,
  sample_metadata = sample_metadata,
  design = ~ group + batch,
  coef = "groupCase",
  level = "celltype_lr"
)

plot_celltype_glm_volcano(glm_res)
```

### Pseudotime module

```{r spatial_pseudotime_intro, eval=FALSE}
# Communication along a pseudotime / time-course (binned, cell-type co-expression).
dyn <- summarize_communication_dynamics(
  reo_demo,
  pseudotime  = pseudotime_vector,
  cell_labels = cell_labels,
  n_bins      = 5
)
plot_communication_dynamics(dyn, level = "pair", metric = "sum_lcs")
```

> **Spatial.** Graph-based spatial scoring (`summarize_spatial_communication()`)
> is pending re-implementation in the v0.12 cell-type rewrite; it currently warns
> and returns cell-type co-expression. `build_spatial_graph()` and
> `plot_spatial_logic()` still work.


## Step 11: Filter and Export

```{r filter_export, eval=FALSE}
# Keep only significant Case-enriched pairs
sig_pairs <- filter_lcs(result$comparison,
                         min_asymmetry = 0.3,
                         max_fdr       = 0.05,
                         direction     = "up")

# Export to CSV
export_results(sig_pairs, file = "LogicComm_significant_pairs.csv")
```

---

## Interpretation Guide

| Output column | Meaning |
|---|---|
| `asymmetry` | Case_freq − Ctrl_freq. Range [−1, +1]. +0.8 means 80% more Case samples active. |
| `case_freq` | Fraction of Case samples with LCS ≥ threshold for this pair |
| `lcs` | Logic Consensus Score: cell-type co-expression (fraction of the sender x receiver opportunity with the L→R logic complete) |
| `shift_score` | Rank shift > 0: gene climbed in transcriptomic dominance in Case |
| `log2fc_lcs` | Traditional fold-change in mean LCS (can be misleading alone) |
| `fdr_fisher` | BH-adjusted p from Fisher's exact test on frequency table |

---

## Current Behavior Notes

- Multimer ligands/receptors require all subunits to be present in the REO matrix and active in the same cell.
- `group_info` must contain one label for every sample.
- In multi-sample comparison, L-R pairs are aligned by name across samples; missing pairs are treated as `NA`.
- `filter_lcs(direction = ...)` accepts only `"up"`, `"down"`, or `"both"`.
- `plot_lcs_bubble(color_by = ...)` accepts only `"log2fc"` or `"asymmetry"`.

---

## Troubleshooting

### Missing dependencies during install

```{r install_missing, eval=FALSE}
install.packages(c("dplyr", "ggplot2", "ggrepel", "scales", "Rcpp"))
```

### Deprecated neighborhood arguments

If you see a warning that `knn_mat` / `mode` / `graph_*` is deprecated and
ignored, drop those arguments: LogicComm scores at the cell-type level from
co-expression and no longer uses a per-cell graph.

### Heatmap dependency

```{r install_pheatmap, eval=FALSE}
install.packages("pheatmap")
```

---

## Session Info

```{r sessioninfo}
sessionInfo()
```
