---
title: "PBMC Multi-Sample Simulation Demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PBMC Multi-Sample Simulation Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  message = FALSE,
  warning = FALSE
)
```

## Purpose

This vignette demonstrates a sample-level Case versus Control validation using the
same PBMC3K dataset that is used in the Seurat tutorial. The goal is not to claim
that PBMC3K contains disease biology. Instead, one PBMC3K object is split into
four pseudo-samples named `pbmc1`, `pbmc2`, `pbmc3`, and `pbmc4`; `pbmc1` and
`pbmc2` are treated as simulated Case samples, while `pbmc3` and `pbmc4` are
left as simulated Controls. A controlled perturbation is added to selected
ligand-receptor genes in the Case samples so that the expected answer is known
before running LogicComm.

**中文说明。** 这个示例把同一个 PBMC3K 数据拆成四个样本级对象：`pbmc1`、`pbmc2`、
`pbmc3` 和 `pbmc4`。其中 `pbmc1`、`pbmc2` 是模拟 Case，`pbmc3`、`pbmc4` 是模拟
Ctrl。我们人为增强 Case 中的 MIF、galectin 和 MHC-I 相关基因，用来校验
`run_multisample()` 和 `CompareLogicGroups()` 是否能在样本层面恢复已知扰动。

Because there are only two simulated samples per group, Fisher-test p-values have
very low resolution. In this demonstration, successful validation is defined by
recovering the expected direction, recurrence, and effect size: targeted LR pairs
should be active in both Case samples, inactive in both Control samples at the
selected threshold, and have positive `log2fc_lcs`. Real publication claims need
more independent biological samples.

## Executable Script

The full executable script is installed with the package at
`inst/examples/pbmc_multisample_demo.R`. After installing LogicComm and the PBMC3K
SeuratData package, run:

```{r run_script, eval=FALSE}
source(system.file("examples", "pbmc_multisample_demo.R", package = "LogicComm"))
demo <- run_pbmc_multisample_demo(
  seed = 20260526,
  perturb_lambda = 50,
  lcs_threshold = 0.2,
  verbose = TRUE
)

demo$target_rows[, c(
  "lr_pair", "pathway", "case_freq", "ctrl_freq", "asymmetry",
  "case_mean_lcs", "ctrl_mean_lcs", "log2fc_lcs", "fdr_fisher"
)]
```

The script performs these operations:

1. Loads PBMC3K with `pbmc3k.SeuratData` and updates the object with Seurat.
2. Extracts the raw RNA count matrix.
3. Randomly splits cells into four sample-level matrices named `pbmc1` to `pbmc4`.
4. Adds a controlled positive perturbation to `MIF`, `CD74`, `LGALS3`, `CD44`,
   `HLA-A`, and `CD8A` in `pbmc1` and `pbmc2` only.
5. Runs `run_multisample()` with the built-in `lr_pairs_human` database.
6. Reports the targeted LR pairs that should be recovered.

## Expected Validation Result

Using LogicComm 0.6.6 on PBMC3K with `seed = 20260526`, `perturb_lambda = 50`,
and `lcs_threshold = 0.2`, the targeted Case perturbation is recovered as follows:

| lr_pair | pathway | case_freq | ctrl_freq | asymmetry | log2fc_lcs | fdr_fisher |
|---|---|---:|---:|---:|---:|---:|
| HLA-A_CD8A | MHC-I | 1 | 0 | 1 | 4.7840 | 1 |
| HLA-B_CD8A | MHC-I | 1 | 0 | 1 | 4.5520 | 1 |
| LGALS3_CD44 | GALECTIN | 1 | 0 | 1 | 4.5161 | 1 |
| MIF_CD74 | MIF | 1 | 0 | 1 | 3.5407 | 1 |
| MIF_CD44 | MIF | 1 | 0 | 1 | 5.3543 | 1 |

The `case_freq = 1` and `ctrl_freq = 0` pattern means that every targeted LR pair
is above the active threshold in both simulated Case PBMC samples and below the
threshold in both simulated Control PBMC samples. The FDR column remains 1 because
with only two samples per group the Fisher test cannot provide publication-level
resolution. This is an intentional validation demo, not a powered disease study.

## Plotting the Comparison

```{r plot_demo, eval=FALSE, fig.width=11, fig.height=5}
comparison <- demo$result$comparison
group_info <- demo$group_info

plot_lcs_bubble(
  comparison,
  top_n = 15,
  min_asymmetry = 0.2,
  color_by = "asymmetry"
)

plot_lcs_heatmap(
  comparison,
  group_info = group_info,
  top_n = 20,
  fdr_cutoff = 1
)
```

For a real cohort, use one PBMC object per biological sample rather than splitting
one sample. Keep the same LR database, REO threshold, layer, and graph definition
across samples. If Seurat nearest-neighbor graphs are available per sample, pass
`graph_name = "RNA_nn"` to `run_multisample()`; otherwise LogicComm falls back to
global LCS mode, which is the mode used by this compact validation script.

## Interpreting This Demo

This demo verifies three package-level behaviors. First, named sample alignment is
preserved from `sample_list` through `group_info` into the comparison table.
Second, REO and LCS scoring respond predictably to a known PBMC-derived
perturbation. Third, `CompareLogicGroups()` reports recurrence and effect size at
the sample level rather than treating cells as independent replicates.

**中文说明。** 这个 demo 的重点是软件校验：样本名是否严格对齐、已知扰动是否被恢复、
多组比较是否以样本为统计单位。它不能替代真实 Case/Ctrl 队列，也不应把 PBMC3K 的模拟
扰动解释成真实疾病机制。真实发表分析至少需要多个独立生物样本，并应结合置换检验、
边支持度、细胞类型解释和外部生物学证据。
