---
title: "Seurat_demo: PBMC3K LogicComm Tutorial / PBMC3K 中英文教程"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Seurat_demo: PBMC3K LogicComm Tutorial / PBMC3K 中英文教程}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  message = FALSE,
  warning = FALSE
)
```

## Overview

> ### ⚠️ v0.12 API update — read first
>
> LogicComm v0.12 scores cell-cell communication at the **cell-type level from
> REO co-expression**; the per-cell **KNN/SNN neighborhood** scoring was removed
> (for dissociated scRNA-seq that graph is in expression space, not physical
> space, so it cannot license spatial juxtacrine/paracrine claims). The
> recommended entry point is now the one-call pipeline:
>
> ```r
> # REO with the rank matrix (only needed for lcs_weighting = "rank").
> reo <- calc_REO_matrix(GetAssayData(pbmc, layer = "counts"),
>                        lr_genes = all_lr_genes(lr_pairs_human), return_rank = TRUE)
> res <- discover_celltype_communication(
>   reo_mat       = reo,
>   cell_labels   = Idents(pbmc),
>   expr          = GetAssayData(pbmc, layer = "counts"),  # enables proliferation filter
>   lcs_weighting = "rank",      # optional REO-intensity score (default "binary")
>   n_perm = 1000, n_cores = 8, seed = 1, fdr_cutoff = 0.1
> )
> res$shortlist   # axes passing per-axis permutation_fdr <= 0.1
> res$view        # confound-filtered discovery view
> ```
>
> The sections below still show the legacy per-step / neighborhood API. Old
> arguments (`knn_mat`, `mode`, `graph_name`, `graph_symmetrize`,
> `edge_weight_mode`, `remove_self_edges`) are **accepted but deprecated and
> ignored with a warning**, so the tutorial still runs; results are scored from
> cell-type co-expression regardless. A full rewrite of the per-step sections to
> the cell-type API is in progress.

This tutorial shows how to run **LogicComm** on the classic **PBMC3K** dataset
with a Seurat workflow.

**中文说明。** 本教程演示如何在经典 **PBMC3K** 数据集上结合 Seurat 使用
**LogicComm**。英文段落保留方法学细节，中文段落提供对应解释，方便中英文对照阅读。

The goal is to demonstrate a practical single-sample workflow:

1. Load PBMC3K as a `Seurat` object.
2. Run standard Seurat preprocessing (clustering/UMAP for cell-type labels).
3. Convert expression to a LogicComm REO matrix.
4. Calculate ligand-receptor Logic Consensus Scores (LCS) from cell-type co-expression.
5. Score per-cell communication activity.
6. Visualize communication activity on UMAP.
7. Optionally summarize top active L-R pairs and cluster-level signals.
8. Summarize celltypeA -> celltypeB communication counts, strengths, pathways,
   and signaling roles.

中文流程对应为：

1. 读取 PBMC3K，并构建 Seurat 对象。
2. 运行标准 Seurat 预处理（聚类/UMAP 用于细胞类型标签）。
3. 将表达矩阵转换成 LogicComm 的 REO 二值逻辑矩阵。
4. 基于细胞类型共表达计算配体-受体 LCS 分数。
5. 评估每个细胞作为发送者或接收者的通讯活性。
6. 在 UMAP 上展示通讯活性。
7. 汇总高活性的配体-受体对和 cluster 层面的信号。
8. 汇总 celltypeA -> celltypeB 的通讯强度、通路和信号角色。

PBMC3K is a single sample, so this tutorial focuses on **within-sample discovery**.
For Case vs Control or cohort-level analysis, see `LogicComm-intro.Rmd`.

**中文说明。** PBMC3K 只有一个样本，因此这里重点展示“单样本内发现”：哪些细胞类型
可能发送信号、接收信号、处于通讯枢纽位置，以及哪些 L-R 轴值得进一步验证。病例-对照
或队列水平比较应使用多个独立生物样本，而不是把所有细胞混在一起当重复。

A publication-style biological question for this vignette is:

> Which PBMC cell types act as communication senders, receivers, mediators, or
> influencers, and which ligand-receptor/pathway axes explain the strongest
> sender -> receiver relationships?

The later sections translate raw LogicComm scores into figure-ready summaries,
role confidence, edge-support QC, optional receiver-response evidence, spatial
neighborhood analysis, pseudotime dynamics, and sample-level differential models.

**中文说明。** 后续章节会把原始 LCS 分数转成更适合发表图表和生物学解释的结果，包括
细胞类型通讯网络、角色分析、边支持度质控、受体下游响应证据、空间邻域扩展、拟时序动态
以及多样本差异模型。

### Key terms / 关键术语

| English term | 中文解释 |
|---|---|
| REO matrix | 相对表达排序得到的二值逻辑矩阵，表示每个细胞内基因是否相对活跃。 |
| LCS | Logic Consensus Score，v0.12 中表示配体-受体逻辑在“发送细胞类型 × 接收细胞类型”共表达机会中被满足的比例。 |
| cell-type co-expression | v0.12 的打分单元：每个 sender→receiver 细胞类型对，配体在 sender、受体在 receiver 同时活跃的比例（不依赖 per-cell 图）。 |
| 已弃用的 neighborhood mode | 旧版基于 Seurat KNN/SNN 的 per-cell 邻域打分；表达空间的图不能代表物理邻近，已移除（相关参数接受但忽略并警告）。 |
| self-loop | 单个细胞自己连自己，例如 `CellA -> CellA`，默认去掉。 |
| same-cell-type signaling | 同一细胞类型内部通讯，例如 `T cell -> T cell`，默认保留。 |
| autocrine-like / homotypic | 自分泌样或同类细胞通讯潜力，不能直接证明同一个单细胞自己分泌并自己接收。 |

---

## Installation

**中文说明。** 安装分为三部分：R 包基础依赖、Seurat 工作流依赖，以及可选可视化依赖。
如果只是检查 LogicComm 包本身，可以先不安装 Seurat；如果要完整运行本教程，需要安装
Seurat 和 PBMC3K 数据。

```{r install, eval=FALSE}
# Core dependencies
install.packages(c(
  "Matrix", "dplyr", "ggplot2", "ggrepel", "scales", "Rcpp",
  "testthat", "rmarkdown"
))

# Seurat workflow dependencies
install.packages(c("Seurat", "SeuratObject"))

# Optional but useful
install.packages("pheatmap")

# Install LogicComm from a local source tarball (adjust the version number)
install.packages("LogicComm_0.9.3.tar.gz", repos = NULL, type = "source")

# Or install a published release directly from GitHub:
# install.packages(
#   "https://github.com/awaragml00029-debug/LogicComm/releases/download/v0.9.3/LogicComm_0.9.3.tar.gz",
#   repos = NULL, type = "source"
# )
```

The PBMC3K data are available from SeuratData:

```{r install_pbmc3k, eval=FALSE}
install.packages("remotes")
remotes::install_github("satijalab/seurat-data")
SeuratData::InstallData("pbmc3k")
```

If you cannot install `SeuratData`, use the manual 10X download shown near the
end of this tutorial.

---

## Load Packages

**中文说明。** 这里加载 LogicComm、Seurat、PBMC3K 数据包、稀疏矩阵和 ggplot2。
如果某些可选包没有安装，只会影响对应扩展图或教程块，不影响核心 LCS 计算。

```{r load_packages, eval=FALSE}
library(LogicComm)
library(Seurat)
library(SeuratData)
library(Matrix)
library(ggplot2)
```

---

## Step 1: Load PBMC3K

**中文说明。** 第一步读取 PBMC3K。它是一个经典外周血单细胞数据集，适合演示
免疫细胞之间的通讯发现流程。

```{r load_pbmc3k, eval=FALSE}
# Load the SeuratData PBMC3K object
SeuratData::data("pbmc3k")
pbmc <- pbmc3k
pbmc
```

PBMC3K contains roughly 2,700 peripheral blood mononuclear cells. The expression
matrix is already stored inside a Seurat object.

**Biological interpretation.** PBMC3K is a healthy peripheral blood immune-cell
mixture, typically containing T cells, B cells, NK cells, CD14 monocytes, FCGR3A
monocytes, dendritic cells, and a small platelet/megakaryocyte-like population.
Because this is not a disease-vs-control dataset, the main biological question is
not differential signaling, but which immune communication programs are active
inside this one resting PBMC sample.

**中文说明。** 因为 PBMC3K 不是疾病组和对照组的比较数据，所以不要把结果解读成
“疾病差异通讯”。更合适的问题是：在这个健康 PBMC 样本中，哪些免疫细胞群体可能在
发送、接收或协调配体-受体信号。

If your object uses Seurat v5, LogicComm can read a named assay layer such as
`"counts"`. In Seurat v4, this corresponds to the counts slot.

---

## Step 2: Standard Seurat Preprocessing

LogicComm itself does not require Seurat normalization, but KNN mode needs a
cell-cell graph. We therefore run a normal Seurat preprocessing workflow to build
neighbors and UMAP.

**中文说明。** LogicComm 的 REO 逻辑本身不依赖 Seurat 标准化，但如果要使用
KNN 邻域模式，就需要 Seurat 的细胞-细胞邻居图。因此这里先运行标准 Seurat 流程，
得到 `RNA_nn` / `RNA_snn` 图和 UMAP。

```{r seurat_preprocess, eval=FALSE}
pbmc <- NormalizeData(pbmc, verbose = FALSE)
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000,
                             verbose = FALSE)
pbmc <- ScaleData(pbmc, verbose = FALSE)
pbmc <- RunPCA(pbmc, npcs = 30, verbose = FALSE)

# Standard Seurat neighbors/clustering (used for cell-type labels, not by LogicComm)
pbmc <- FindNeighbors(pbmc, dims = 1:20, verbose = FALSE)

# UMAP for visualization
pbmc <- RunUMAP(pbmc, dims = 1:20, verbose = FALSE)

# Optional clustering for interpretation
pbmc <- FindClusters(pbmc, resolution = 0.5, verbose = FALSE)
```

Seurat usually stores the graph as something like `RNA_snn` or `RNA_nn` in
`pbmc@graphs`. You can inspect available graph names:

```{r inspect_graphs, eval=FALSE}
names(pbmc@graphs)
```

**Biological interpretation.** PCA/UMAP/clustering should separate PBMC immune
states according to marker-gene programs: T/NK cells are often driven by T-cell
and cytotoxic markers, B cells by B-cell receptor markers, monocytes by myeloid
and inflammatory genes, and platelets by megakaryocyte genes. LogicComm later
uses the `RNA_nn` graph as a local cellular neighborhood. Therefore an L-R pair
with high neighborhood LCS means the ligand-active and receptor-active cells are
close in the PBMC transcriptional manifold, not merely present somewhere in the
sample.

**中文说明。** `RNA_nn` 在这里代表转录组空间里的局部邻近关系。高 neighborhood
LCS 的含义不是“这个样本里有 ligand 也有 receptor”这么简单，而是 ligand 活跃细胞
和 receptor 活跃细胞经常出现在邻近关系中。

---

## Step 3: Build the LogicComm REO Matrix

REO converts each cell into binary gene logic states. A gene is active if its
expression is above a within-cell quantile anchor.

**中文说明。** REO 不是普通归一化表达矩阵，而是把每个细胞内的基因表达转换成
“活跃/不活跃”的逻辑状态。这样可以减少测序深度等绝对表达量差异的影响，更关注
每个细胞内部的相对表达排序。

For PBMC3K, we usually only need genes that appear in the built-in
ligand-receptor database. `lr_pairs_human` is a compact 109-interaction human
seed database compiled from public resources including CellChatDB, CellPhoneDB,
and NicheNet, then converted to LogicComm's `ligand_genes` / `receptor_genes`
schema. CellChat's original database is larger and is distributed in the
CellChat repository as `data/CellChatDB.human.rda`; use a custom LR table if you
want to score the full CellChatDB instead of LogicComm's compact seed set:

```{r cellchatdb_convert, eval=FALSE}
data(CellChatDB.human, package = "CellChat")
cellchat_lr <- as_logiccomm_lr_db_from_cellchat(
  CellChatDB.human,
  include_modulators = TRUE
)
cellchat_genes <- all_lr_genes(cellchat_lr, include_modulators = TRUE)
```

Yes, the converted CellChatDB can be used directly as `lr_db` in LogicComm, as
long as you also build the REO matrix with genes from the same converted database:

```{r cellchatdb_direct_use, eval=FALSE}
cellchat_genes <- all_lr_genes(cellchat_lr, include_modulators = TRUE)

reo_pbmc_cellchat <- calc_REO_matrix(
  pbmc,
  lr_genes       = cellchat_genes,
  rank_threshold = 0.5,
  layer          = "counts",
  chunk_size     = 1000,
  return_rank    = TRUE
)

lcs_pbmc_cellchat <- IdentifyLogicConsensus(
  reo_pbmc_cellchat$logic,
  lr_db = cellchat_lr
)

mod_summary <- summarize_lr_modulators(
  reo_pbmc_cellchat$logic,
  cellchat_lr,
  rank_mat = reo_pbmc_cellchat$rank
)

gate_pbmc_cellchat <- IdentifyLogicGateConsensus(
  reo_pbmc_cellchat,
  lr_db          = cellchat_lr,
  positive_gate  = "all",
  negative_gate  = "rank_block"
)

head(gate_pbmc_cellchat[order(gate_pbmc_cellchat$regulated_lcs, decreasing = TRUE), ], 10)
```

Practical cautions when using the full CellChatDB:

**中文说明。** 可以把 CellChatDB 转成 LogicComm 的 LR 数据库格式直接使用，但需要注意：
数据库越大，搜索空间越大，计算和解释都更复杂；不要把每一个正分数都当成已经验证的
生物学事件。

- It is much larger than `lr_pairs_human`, so REO and LCS scoring can take longer.
- More tested interactions also means a larger multiple-testing/search space; rank
  and interpret the top pathways carefully rather than treating every positive
  score as a validated biological event.
- Some CellChat interactions involve complexes or cofactors. The converter expands
  ligand/receptor complexes into subunit lists and can preserve agonist,
  antagonist, co-activating receptor, and co-inhibitory receptor fields. The base
  `IdentifyLogicConsensus()` score still uses only ligand/receptor logic for
  backward compatibility. The enhanced `IdentifyLogicGateConsensus()` translates
  modulators into REO logic gates: positive modulators use the selected
  `positive_gate` mode (`"all"` or `"any"`), while antagonists or inhibitory
  co-receptors can block a candidate event when their
  within-cell rank overtakes the ligand or receptor.
- Biologically, `base_lcs` asks whether L and R are locally co-active, whereas
  `regulated_lcs` asks whether that communication remains valid after CellChatDB
  regulatory gates. A large negative `delta_lcs` or high `block_rate` suggests a
  possible antagonist/co-inhibitory rank reversal: L and R may still be present,
  but the effective communication logic is switched off.
- Use one LR database consistently across all samples if you plan to compare
  samples or groups.

```{r calc_reo, eval=FALSE}
data(lr_pairs_human)

lr_genes <- all_lr_genes(lr_pairs_human)

reo_pbmc <- calc_REO_matrix(
  pbmc,
  lr_genes        = lr_genes,
  rank_threshold  = 0.5,
  layer           = "counts",
  chunk_size      = 2000,
  gene_background = "quantile",   # cell-type-free ambient guard (recommended)
  verbose         = TRUE
)

reo_pbmc
```

**Why `gene_background = "quantile"`?** The within-cell REO anchor asks only
"is this gene above this cell's own median gene", so a marker like `CD8A` that
leaks into non-CD8 cells as ambient RNA (a few counts, above the cell's low
median) is wrongly called active — producing biologically impossible axes such
as `HLA-B -> CD8A` with a Treg receiver. The `gene_background = "quantile"`
guard additionally requires each gene to exceed its own background, taken over
the cells that *detect* it (nonzero counts) so it stays effective even for
markers expressed in a minority of cells; this removes ambient activity (it
needs no cell-type labels). For rigorous ambient removal, run SoupX, DecontX, or
CellBender upstream and then build the REO matrix on the corrected counts.

Important notes:

- Input is interpreted as **genes x cells**.
- For Seurat objects, LogicComm extracts the assay matrix internally.
- Sparse matrices are supported.
- LogicComm uses all genes for REO anchors by default, but processes only retained
  output genes by cell chunks to reduce memory pressure.
- `lr_genes` controls the output/scored genes, not the background anchor universe.
  Use `anchor_genes` only for deliberate sensitivity analyses.
- `rank_threshold = 0.5` uses the within-cell median as the activity anchor.
  Increase it, for example to `0.6` or `0.75`, for stricter activity calls.

**Biological interpretation.** The REO matrix is not another normalized
expression matrix. It is a binary map of whether each LR gene is relatively active
inside each individual cell. In PBMCs, this helps reduce dependence on absolute
library size and emphasizes logic states such as "this monocyte has IL1B active"
or "this T cell has a chemokine receptor active". A stricter threshold selects
only stronger within-cell activity and may reduce noisy low-level immune signals.

---

## Step 4: Compute Logic Consensus Scores (LCS)

`IdentifyLogicConsensus()` computes, for each L-R pair, a **dataset-level** Logic
Consensus Score: the fraction of cells in which the complete ligand logic and the
complete receptor logic are simultaneously realized (co-expression). LogicComm
v0.12 scores from REO co-expression and no longer uses a per-cell KNN/SNN graph.

**中文说明。** 这一步计算每个 L-R 对的 LCS：在多大比例的细胞里配体逻辑和受体逻辑
同时成立（共表达）。v0.12 起 LogicComm 基于 REO 共表达打分，不再使用 per-cell 邻域图。

```{r lcs_global, eval=FALSE}
lcs_pbmc <- IdentifyLogicConsensus(
  reo_pbmc,
  lr_db         = lr_pairs_human,
  lcs_threshold = 0.01,
  verbose       = TRUE
)

lcs_pbmc
```

View the top active L-R pairs:

```{r top_lcs, eval=FALSE}
top_lcs <- sort(lcs_pbmc[!is.na(lcs_pbmc)], decreasing = TRUE)
head(top_lcs, 20)
```

**Biological interpretation.** A high LCS means the ligand and receptor are
frequently co-expressed across the sample. In PBMC3K, top pairs often reflect
common immune biology rather than disease-specific effects: chemokine axes may
indicate immune-cell positioning; cytokine axes may reflect inflammatory
readiness; integrin/collagen/ECM-like axes can reflect adhesion programs;
checkpoint or co-stimulatory axes can indicate lymphocyte regulatory states.
Treat the top list as hypotheses; confirm cell types with marker genes before a
strong claim.

> The dataset-level `IdentifyLogicConsensus()` score measures co-expression
> across all cells (most informative for autocrine-like / same-cell co-expression).
> For directional **sender -> receiver cell-type** communication, use
> `summarize_celltype_communication()` (Step 9) or the one-call
> `discover_celltype_communication()`.

> **Migration note.** Legacy neighborhood arguments (`seurat_obj`, `graph_name`,
> `knn_mat`, `remove_self_edges`, `graph_symmetrize`, `edge_weight_mode`) are
> accepted but deprecated and ignored with a warning.

### 4.1 Same-cell-type signaling and autocrine-like interpretation

LogicComm follows the same practical interpretation used by many cell-cell
communication tools: a same-cell-type edge such as `Tumor -> Tumor` or
`Macrophage -> Macrophage` is kept by default and interpreted as same-population,
homotypic, or autocrine-like signaling potential. This result does not prove that
the exact same single cell secreted and received the ligand. It means the sender
and receiver populations are the same annotated cell type.

**中文说明。** LogicComm 默认保留同一细胞类型内部通讯，这和 CellChat 中常见的
`celltype -> same celltype` 解释方式一致。也就是说，`Tumor -> Tumor` 可以作为
自分泌样或同类细胞通讯潜力来解释，但不要过度声称它证明了“同一个单细胞自己分泌、
自己接收”。默认去掉 `CellA -> CellA` 自环，是为了避免把单细胞内 L/R 共表达直接当成
邻域通讯主证据；默认保留 `celltypeA -> celltypeA`，是为了保留同类细胞群体内部的
通讯信息。

**Biological interpretation.** `RNA_nn` and `RNA_snn` answer slightly different
questions. `RNA_nn` is closer to "which cells are immediate transcriptional
neighbors?" and is therefore the preferred graph for local L-R logic. `RNA_snn`
emphasizes shared neighborhoods and is more cluster-like. If the correlation
between `RNA_nn` and `RNA_snn` LCS is high, the strongest signals are robust to
the neighborhood definition. If it is low, interpret graph-dependent pairs more
cautiously and inspect whether they are driven by cluster boundaries.

---

## Step 5: Convert LCS Vector to a Table

**中文说明。** `IdentifyLogicConsensus()` 返回的是一个命名向量。为了查看、排序、保存
或和其他结果合并，通常先把它转成普通表格。

```{r lcs_table, eval=FALSE}
lcs_tbl <- data.frame(
  lr_pair = names(lcs_pbmc),
  lcs     = as.numeric(lcs_pbmc),
  stringsAsFactors = FALSE
)

lcs_tbl <- lcs_tbl[!is.na(lcs_tbl$lcs), ]
lcs_tbl <- lcs_tbl[order(lcs_tbl$lcs, decreasing = TRUE), ]

head(lcs_tbl, 20)
```

**Biological interpretation.** The table is the ranked communication-program
summary for this sample. The `lcs` value is useful for prioritization: higher
values indicate more frequent neighborhood-level ligand/receptor co-activity.
Rows near zero are not necessarily biologically impossible; they may be absent in
PBMC3K, too rare, filtered by the REO threshold, or require a cell type not well
represented in the dataset. For publication-style interpretation, annotate top
rows by pathway and inspect whether their ligand and receptor genes are expressed
in plausible PBMC populations.

You can save the table for downstream analysis:

```{r export_lcs, eval=FALSE}
write.csv(lcs_tbl, "PBMC3K_LogicComm_LCS.csv", row.names = FALSE)
```

---

## Step 6: Score Per-Cell Communication Activity

`score_lr_activity()` estimates how strongly each cell participates in active
ligand-receptor logic as sender and/or receiver.

**中文说明。** 这里不是在判断某个具体 L-R 对，而是在问每个细胞总体上更像“发送者”、
“接收者”，还是两者都强。它适合在 UMAP 上快速定位通讯活跃的细胞区域。

```{r score_cells, eval=FALSE}
cell_scores <- score_lr_activity(
  reo_pbmc,
  lr_db     = lr_pairs_human,
  mode      = "both",
  aggregate = TRUE,
  verbose   = TRUE
)

head(cell_scores)
summary(cell_scores$comm_score)
```

Rank the most communication-active cells:

```{r rank_cells, eval=FALSE}
top_cells <- rank_comm_cells(cell_scores, n = 20)
head(top_cells, 20)
```

**Biological interpretation.** `sender_score` highlights cells with many active
ligand programs; `receiver_score` highlights cells with many active receptor
programs; `comm_score` combines both. In PBMC3K, high sender scores may appear in
monocytes or dendritic-like cells because these cells often express cytokines,
chemokines, and inflammatory mediators. High receiver scores may appear in
lymphocytes or monocytes that express chemokine/cytokine receptors. A high score
does not mean the cell is "more important" in every biological sense; it means it
matches more active LR logic patterns under the selected LR database and REO
threshold.

---

## Step 7: Add Scores Back to Seurat Metadata

**中文说明。** 把 LogicComm 分数写回 Seurat metadata 后，就可以直接用 Seurat 的
`FeaturePlot()`、`VlnPlot()` 等函数展示通讯活性。

```{r add_metadata, eval=FALSE}
# score_lr_activity() returns a list, not a data.frame.
# Align each named score vector to the Seurat cell order before adding metadata.
cell_order <- colnames(pbmc)

pbmc$LogicComm_sender   <- cell_scores$sender_score[cell_order]
pbmc$LogicComm_receiver <- cell_scores$receiver_score[cell_order]
pbmc$LogicComm_score    <- cell_scores$comm_score[cell_order]
```

Plot communication activity on UMAP:

```{r umap_logiccomm, eval=FALSE, fig.width=10, fig.height=4}
FeaturePlot(
  pbmc,
  features  = c("LogicComm_sender", "LogicComm_receiver", "LogicComm_score"),
  reduction = "umap",
  ncol      = 3
)
```

You can also compare by Seurat cluster:

```{r violin_cluster, eval=FALSE, fig.width=8, fig.height=4}
VlnPlot(
  pbmc,
  features = "LogicComm_score",
  group.by = "seurat_clusters",
  pt.size  = 0
)
```

**Biological interpretation.** The UMAP shows where communication-active cells
sit in the immune-cell landscape. If sender activity localizes to a monocyte
region, that suggests myeloid cells may be major ligand producers. If receiver
activity localizes to T/NK or B-cell regions, that suggests lymphocytes may be
responding to cytokine/chemokine/adhesion cues. The violin plot asks whether
entire clusters differ in communication potential. Clusters with broad high
scores may represent cell states with many active LR programs, whereas a few high
outlier cells may indicate rare activated cells rather than a whole-cell-type
program.

---

## Step 8: Annotate Top L-R Pairs by Cluster

If you have cluster labels, LogicComm can help identify which clusters show the
logic state for selected L-R pairs.

**中文说明。** 这一步把排名靠前的 L-R 对映射回 cluster 或 cell type，帮助判断可能的
发送细胞群和接收细胞群。它是从“分数列表”走向“生物学解释”的关键步骤。

```{r cluster_labels, eval=FALSE}
# Choose an annotation column if present; otherwise use Seurat clusters.
label_col <- if ("seurat_annotations" %in% colnames(pbmc@meta.data)) {
  "seurat_annotations"
} else {
  "seurat_clusters"
}

cluster_labels <- setNames(
  as.character(pbmc[[label_col]][, 1]),
  colnames(pbmc)
)

# Drop unannotated cells before cluster-level summary.
cluster_labels <- cluster_labels[!is.na(cluster_labels) & nzchar(cluster_labels)]

selected_pairs <- names(head(top_lcs, 10))

cluster_logic <- AutoLabelLogicClusters(
  reo_pbmc[, names(cluster_labels), drop = FALSE],
  cell_labels = cluster_labels,
  lr_pairs    = selected_pairs,
  lr_db       = lr_pairs_human
)

cluster_logic
```

**Biological interpretation.** `cluster_logic` is a bridge between the ranked LR
list and cell-type biology. For each selected pair, it helps identify which
cluster is most consistent with the ligand-active sender state and which cluster
is most consistent with the receptor-active receiver state. For example, a
monocyte-enriched sender cluster paired with a lymphocyte receiver cluster would
support a myeloid-to-lymphoid communication hypothesis. If both sender and
receiver map to the same broad cluster, the signal may represent autocrine-like or
within-lineage communication. Always verify with marker genes and the underlying
ligand/receptor expression patterns.

This is useful for a quick biological interpretation step, especially before
moving to manually curated cell type annotations.


---

## Step 9: Biology-Oriented Cell-Type Communication & Range Analysis

A critical biological question is: **how** do these cells talk?
-   **Juxtacrine**: Required physical proximity (contact-dependent).
-   **Paracrine**: Occurs between neighbors (secreted signaling).
-   **Distal/Endocrine**: Occurs between distant cell types (systemic signaling).

LogicComm now automatically classifies communication events into these ranges by comparing neighborhood-based scores with global potential.

**中文说明。** 这一节是发表级解释的重点：不只看某个 L-R 对有没有分数，还要看它是
局部邻域支持的信号、同类细胞内部信号，还是可能被 KNN 图漏掉的远距离候选信号。
LogicComm 会同时报告 local neighborhood evidence 和 global candidate potential，
避免把远距离通讯简单丢掉。

```{r summarize_celltype_comm, eval=FALSE}
# Use a biological annotation column
label_col <- if ("seurat_annotations" %in% colnames(pbmc@meta.data)) {
  "seurat_annotations"
} else {
  "seurat_clusters"
}

celltype_labels <- setNames(as.character(pbmc[[label_col]][, 1]), colnames(pbmc))

ct_comm <- summarize_celltype_communication(
  reo_mat       = reo_pbmc,
  cell_labels   = celltype_labels,   # no graph: scored from cell-type co-expression
  include_self  = TRUE,              # keep same-cell-type pairs, e.g. T cell -> T cell
  lr_db         = lr_pairs_human,
  lcs_threshold = 0.01,
  min_edges     = 20,                # min sender x receiver cell-count opportunity
  verbose       = TRUE
)
```

### 9.1 A note on communication "range"

v0.12 scores at the **cell-type level from co-expression**, so the former
local / distal "communication range" — a per-cell neighborhood concept — no
longer applies: every axis is a cell-type co-expression axis. Genuine
distance/range claims require physical spatial coordinates (see the spatial
module). Pathway- and strength-level overviews are in §9.4–§9.5.

**中文说明。** v0.12 在细胞类型层面用共表达打分，旧的 local/distal“通讯距离”属于
per-cell 邻域概念，已不再适用——每条都是细胞类型共表达 axis。真正的距离声称需要物理
空间坐标（见空间模块）。

### 9.2 Publication-Ready Network Visualization
Visualize the sender -> receiver cell-type communication network.

```{r celltype_network, eval=FALSE, fig.width=8, fig.height=8}
plot_celltype_network(
  ct_comm,
  metric = "sum_lcs",
  layout = "circle"
)
```

### 9.3 Signaling Roles & Hubs
Identify which cell types drive the communication network.

**中文说明。** 角色分析回答的是：哪些细胞类型更像发送者，哪些更像接收者，哪些可能是
网络中的 hub、mediator 或 influencer。角色标签需要结合 evidence 和 confidence 一起看，
不要只看一个最高分。

```{r signaling_roles, eval=FALSE, fig.width=10, fig.height=5}
# Outgoing vs Incoming strength landscape
plot_celltype_roles(ct_comm)

# Detailed role metrics heatmap
plot_celltype_role_heatmap(ct_comm)
```

### 9.4 Mechanistic L-R Hotspots
Inspect specific L-R pairs driving the communication between cell types.

```{r lr_bubble_advanced, eval=FALSE, fig.width=10, fig.height=7}
# Advanced bubble plot with range classification and pathway faceting
plot_lr_bubble_advanced(
  ct_comm,
  top_n_per_pair = 5
)
```

**Interpretation.** This plot allows you to see the "Logic Hotspots" across the entire immune landscape. Faceting by pathway helps separate functional modules (e.g., Cytokines vs. ECM), while bubble color confirms if the interaction is local or distal.

**中文说明。** bubble 图适合快速找出细胞类型之间最突出的 L-R 热点。分面可以区分不同
功能模块，颜色可以提示该信号主要来自局部邻域证据还是远距离/全局候选证据。

### 9.5 Pathway-Level Dominance
Summarize pathway-level communication strength and network structure.

```{r pathway_dominance_network, eval=FALSE, fig.width=10, fig.height=7}
plot_pathway_heatmap(
  ct_comm,
  metric = "sum_lcs",
  top_n_pathways = 25
)

plot_pathway_dominance(ct_comm)
plot_pathway_network(ct_comm, pathway = "MIF", top_n_edges = 20)
```

In PBMC data, MHC-I, MHC-II, and MIF can dominate the total communication
landscape. That is biologically meaningful, but not always a pair-specific
mechanism. The next step is to separate stable but broad immune-recognition axes
from more specific sender -> receiver candidates.

**中文说明。** 在 PBMC 中，MHC 和 MIF 等通路可能非常强，但强并不等于一定是最特异的
机制。发表时最好区分“广泛免疫识别背景信号”和“特定 sender -> receiver 的机制候选”。

```{r communication_specificity, eval=FALSE}
ct_comm <- score_communication_specificity(ct_comm)
head(ct_comm$specificity_summary[order(-ct_comm$specificity_summary$total_lcs), ], 20)
```

For one selected event, inspect numerical evidence and optionally draw a UMAP
logic plot:

```{r explain_celltype_event, eval=FALSE, fig.width=7, fig.height=5}
# Replace lr_pair/sender/receiver with a row that exists in ct_comm$lr_table.
plot_lr_evidence(
  ct_comm,
  sender   = "CD14 Mono",
  receiver = "CD4 T / mixed T",
  lr_pair  = "CXCL12_CXCR4"
)

ex <- explain_celltype_interaction(
  ct_comm,
  sender     = "CD14 Mono",
  receiver   = "CD4 T / mixed T",
  lr_pair    = "CXCL12_CXCR4",
  reo_mat    = reo_pbmc,
  seurat_obj = pbmc
)

ex$evidence
ex$plot
```

### 9.4 Audit a significant sender-receiver pair

After a permutation result such as `head(null_pair[order(null_pair$fdr), ], 10)`,
the next question is not only whether a sender-receiver pair is significant. For
publication interpretation, audit the pair by asking which L-R pairs drive it,
whether those L-R pairs are broad or pair-specific, whether edge support is
adequate, whether specificity is high, and whether the signal repeats across
biological samples.

**中文说明。** `null_pair` 显著只能说明某个 sender -> receiver 组合高于随机标签背景。
发表解释时还要继续追问：到底哪些 L-R pair 在驱动？是不是 MHC/MIF 这类广泛轴？边支持
是否足够？specificity 是否高？在多样本中是否重复出现？

```{r audit_significant_celltype_pair, eval=FALSE}
# Choose the strongest permutation-supported sender-receiver pair.
top_pair <- null_pair[order(null_pair$fdr), ][1, ]
audit_sender <- top_pair$sender_type
audit_receiver <- top_pair$receiver_type

# Add specificity annotations before auditing L-R drivers.
ct_comm <- score_communication_specificity(ct_comm)

lr_audit <- ct_comm$lr_table[
  ct_comm$lr_table$sender_type == audit_sender &
    ct_comm$lr_table$receiver_type == audit_receiver &
    ct_comm$lr_table$active %in% TRUE &
    is.finite(ct_comm$lr_table$lcs),
  ,
  drop = FALSE
]

lr_audit <- lr_audit[order(-lr_audit$lcs, -lr_audit$n_active_edges), ]

audit_cols <- intersect(c(
  "sender_type", "receiver_type", "lr_pair", "pathway",
  "lcs", "n_edges", "n_active_edges", "n_global_possible", "n_active_global",
  "sender_cell_count", "receiver_cell_count",
  "ligand_active_frac_sender", "receptor_active_frac_receiver",
  "pair_specificity", "specificity_class"
), names(lr_audit))

head(lr_audit[, audit_cols, drop = FALSE], 20)
```

Interpret the audit table as follows:

1. **L-R drivers.** The rows with the largest `lcs` and `n_active_edges` are the
   main L-R pairs driving this sender-receiver signal.
2. **Broad or pair-specific.** `specificity_class` separates
   `Pair-specific candidate` from `Broad/ubiquitous axis` and
   `Identity-associated axis`. Broad axes can be real biology, but should not be
   overclaimed as a highly specific mechanism.
3. **Opportunity support.** `n_edges` / `n_global_possible` is the sender x
   receiver cell-count opportunity, and `n_active_edges` / `n_active_global` is
   how much of it co-expresses the L-R pair. A high `lcs` resting on very few
   cells (small `sender_cell_count` / `receiver_cell_count`) is less robust than
   the same `lcs` over abundant populations.
4. **Specificity score.** `pair_specificity` close to 1 means the L-R pair is
   concentrated in fewer sender-receiver combinations. Low values mean the L-R
   pair is distributed broadly across the cell-type network.
5. **Multi-sample recurrence.** A single PBMC object can support mechanism
   discovery, but publication-level claims require recurrence across biological
   samples.

If you have already built `ct_comparison` from multiple samples, check whether
the same celltypeA -> celltypeB -> L-R feature is recurrent:

```{r audit_multisample_recurrence, eval=FALSE}
top_lr <- lr_audit$lr_pair[1]
feature_id <- paste(audit_sender, audit_receiver, top_lr, sep = "|")

ct_comparison[ct_comparison$lr_pair == feature_id, intersect(c(
  "lr_pair", "case_freq", "ctrl_freq", "asymmetry",
  "case_mean_lcs", "ctrl_mean_lcs", "log2fc_lcs",
  "n_case_avail", "n_ctrl_avail", "fdr_fisher"
), names(ct_comparison)), drop = FALSE]
```

If this row is absent or has low `n_case_avail` / `n_ctrl_avail`, the feature has
not yet passed the sample-recurrence check. In that case, describe it as a
single-dataset candidate and validate it in an independent cohort.

### 9.5 Receiver downstream response evidence

L-R logic identifies communication potential. A stronger biological argument asks
whether the receiver also shows downstream response-state logic. You can provide
a lightweight response database with `lr_pair` and `response_genes` (semicolon,
comma, pipe, or plus separated):

**中文说明。** L-R 共活跃只能说明“有通讯潜力”。如果接收细胞同时出现下游响应基因的
逻辑活性，生物学证据会更强。这个模块就是把 receiver response 作为额外证据整合进去。

```{r receiver_response_score, eval=FALSE}
response_db <- data.frame(
  lr_pair = c("IFNG_IFNGR1_IFNGR2", "TGFB1_TGFBR1_TGFBR2"),
  response_genes = c("STAT1;IRF1;CXCL10", "SMAD7;SERPINE1;COL1A1"),
  stringsAsFactors = FALSE
)

ct_comm <- add_receiver_response_score(
  ct_comm,
  reo_mat       = reo_pbmc,
  response_db   = response_db,
  response_mode = "any"
)

head(ct_comm$lr_table[order(ct_comm$lr_table$response_integrated_score, decreasing = TRUE), ], 10)
```

### 9.6 Bootstrap, permutation null, and threshold sensitivity

For publication-level interpretation, inspect uncertainty and null expectations.
The bootstrap is fast and approximate; the permutation null is slower but asks
whether the observed sender-receiver structure exceeds random cell-type labels.

**中文说明。** 发表级分析不能只展示最高分，还需要回答三个问题：这个信号是否稳定？
是否高于随机标签背景？是否对 REO 阈值敏感？bootstrap、permutation null 和阈值敏感性
分别对应这三个检查。

```{r celltype_bootstrap, eval=FALSE}
boot_pair <- bootstrap_celltype_communication(
  ct_comm,
  n_boot = 200,
  level  = "celltype_pair",
  seed   = 1
)
head(boot_pair)
```

The permutation null shuffles cell-type labels to ask whether each
sender -> receiver pair is stronger than chance. Two practical points:

- **Permutations are independent, so set `n_cores`** to fork the loop (Unix/macOS)
  for a near-linear speedup that makes a large `n_perm` affordable.
- **The empirical p-value floor is `1 / (n_perm + 1)`.** With `n_perm = 12` the
  smallest possible p is 0.083, so nothing reaches p < 0.05 and no axis can earn
  Tier 1. Use `n_perm >= 200` (floor ~0.005) for single-axis resolution, or
  `n_perm = 1000` to leave headroom for BH correction across many axes.

```{r celltype_permutation_null, eval=FALSE}
null_pair <- permute_celltype_communication(
  ct_comm = ct_comm,
  reo_mat = reo_pbmc,
  n_perm  = 1000,           # axis-level null (default metric = "lcs"); floor ~1/1001
  n_cores = 8,              # fork-parallel on Unix/macOS: ~Nx speedup
  seed    = 1
)
# Axis-level: every sender -> receiver -> L-R axis gets its OWN empirical p and
# BH FDR (null_pair has an lr_pair column); rank_communication_axes() joins them
# per axis and exposes permutation_fdr. The co-expression null needs no graph.
head(null_pair[order(null_pair$fdr), ], 10)

diagnose_permutation_resolution(null_pair)

# If null_sd == 0 and observed > null_mean, LogicComm reports z_score = Inf and
# degenerate_positive_null = TRUE. This means shuffled labels never created a
# comparable edge. Biologically, treat it as a structural-null signal that needs
# edge support, specificity, and independent validation, not as an ordinary
# Gaussian z-score.
null_pair[null_pair$degenerate_positive_null %in% TRUE, c(
  "sender_type", "receiver_type", "observed", "null_mean", "z_score",
  "empirical_p", "fdr", "null_biological_interpretation"
)]
```

```{r reo_threshold_sensitivity, eval=FALSE}
sens <- sensitivity_REO_threshold(
  pbmc,
  rank_threshold_grid = c(0.4, 0.5, 0.6, 0.7),
  lr_db       = lr_pairs_human,
  cell_labels = celltype_labels,
  layer       = "counts"
)
head(sens$stability, 20)

plot_specificity_stability(sens, ct_comm = ct_comm)

# Flag the proliferation / transcriptional-breadth confound BEFORE ranking, so the
# hub flags ride along into ranked_axes (diagnostic only; see Section 9.11). expr
# is a genes x cells matrix aligned to the cells used to build ct_comm.
expr_pbmc <- Seurat::GetAssayData(pbmc, assay = "RNA", layer = "data") # Seurat v4: slot = "data"
ct_comm <- diagnose_proliferation_confound(ct_comm, expr_pbmc)

# Since 0.9.3 rank_communication_axes() down-ranks broad/ubiquitous axes by default
# (demote_broad = TRUE): see Section 9.11 for why and how to read the result.
ranked_axes <- rank_communication_axes(ct_comm, null_pair = null_pair, sens = sens)
head(ranked_axes, 20)
```

In the stability-specificity plot, the most publication-worthy mechanistic
hypotheses are usually high on both axes. High stability but low specificity is
also important, but it should often be described as a broad immune-recognition or
identity-associated program rather than a novel pair-specific signal.

**Interpretation caution.** The sender -> receiver direction reflects the ligand
side and receptor side of an L-R pair. If your graph is undirected or
symmetrized, both directions may exist as graph opportunities, but the L-R
interpretation remains directional. Very small cell types may have too few edges;
always inspect `n_edges`, adjust `min_edges`, and examine bootstrap/permutation
stability.

**中文说明。** sender -> receiver 的方向来自 ligand 在 sender 侧、receptor 在 receiver
侧。即使 KNN 图被无向化，L-R 的解释仍然是有方向的。小细胞群可能边数不足，所以要同时
检查 `n_edges`、`min_edges`、bootstrap 和 permutation 结果。

---


### 9.7 Publication-ready QC and report generation

A high LCS is more convincing when it is supported by enough edge opportunities,
active edge evidence, and stable role assignments. The following helpers provide
a compact publication-oriented audit trail.

```{r celltype_qc_report, eval=FALSE}
plot_communication_qc(ct_comm)
plot_role_confidence(ct_comm)
plot_celltype_role_dotplot(ct_comm)
plot_pathway_dominance(ct_comm)

diag <- diagnose_celltype_communication(ct_comm, null_pair = null_pair, sens = sens)
diag

findings <- summarize_communication_findings(ct_comm, top_n = 15)
findings$qc
findings$top_celltype_pairs
findings$top_lr_events
findings$role_summary

write_communication_report(
  ct_comm,
  file  = "PBMC3K_LogicComm_report.md",
  title = "PBMC3K LogicComm communication report",
  top_n = 15
)
```

**How to read this for a manuscript.** Use the report as a structured checklist:
first identify the strongest cell-type pairs, then inspect top L-R axes and
pathways, then verify whether the responsible cell types have high-confidence
roles, and finally check whether the signal is supported by enough edges. This
reduces the chance of building a biological story from a rare cell-type pair or a
single unstable L-R event.


#### PBMC-specific interpretation template

When using the compact built-in PBMC-oriented LR seed set, you may see strong
MIF, MHC-I, MHC-II, and GALECTIN contributions. These are plausible immune
programs, but they answer different biological questions:

- **MHC/HLA axes** are often highly stable across REO thresholds. In PBMC, this
  can reflect antigen-presentation or lineage-recognition biology and should be
  reported as a broad immune-recognition / identity-associated program unless it
  also has high sender-receiver specificity or receiver-response evidence.
- **MIF-dominated edges** can identify inflammatory or myeloid-associated
  communication structure, but should still be checked for pathway dominance and
  specificity.
- A cell type such as B cells can be labeled `Mediator (betweenness)` because it
  lies between active communication modules. If `communication_evidence_label` is
  only `Moderate`, write this as **candidate mediator-like bridge** rather than a
  definitive gatekeeper.
- A zero-variance null such as a rare `Platelet -> Platelet` edge should be
  treated as a **structural-null observation**: shuffled labels never reproduced
  the observed score, but small cell abundance and low edge support may still
  make the biological claim weak.

A concise results sentence could be:

> LogicComm identified a PBMC communication landscape dominated by MIF and
> MHC-related immune-recognition programs. Role analysis suggested DC/monocyte
> and lymphoid subsets occupy distinct sender/receiver/mediator tendencies, but
> final biological prioritization favored axes that were not only strong and
> stable, but also cell-type-pair-specific, supported by sufficient edge
> opportunities, and enriched over label-permutation nulls.

### 9.7 Spatial transcriptomics extension

With genuine spatial coordinates (Visium, MERFISH, Xenium, CosMx) a *physical*
neighborhood graph is a legitimate, non-expression-space basis for local
juxtacrine/paracrine scoring — unlike a transcriptional KNN graph.

> **Status (v0.12).** Graph-based spatial **scoring**
> (`summarize_spatial_communication()`) is pending re-implementation in the
> cell-type-co-expression rewrite: it currently warns and returns cell-type
> co-expression (the graph is not used). Graph construction
> (`build_spatial_graph()`) and per-spot REO visualization (`plot_spatial_logic()`)
> still work.

**中文说明。** 用真实空间坐标时，物理邻居图是合理的（不同于表达空间的 KNN）。但 v0.12
的空间**打分**正在按细胞类型重写中暂不可用，目前会警告并返回细胞类型共表达；
`build_spatial_graph()` 和 `plot_spatial_logic()` 仍可用。

```{r spatial_extension, eval=FALSE}
# coords: rownames matching cells/spots; x/y, imagecol/imagerow, or
# pxl_col_in_fullres/pxl_row_in_fullres are recognized automatically.
coords <- spatial_coordinates

sp_graph <- build_spatial_graph(coords, mode = "knn", k = 6, distance_weight = "gaussian")

plot_spatial_logic(
  reo_pbmc,
  coords,
  lr_pair     = "VEGFA_KDR",
  lr_db       = lr_pairs_human,
  cell_labels = celltype_labels
)
```

### 9.8 Pseudotime and time-course communication dynamics

For differentiation or activation trajectories, communication may rewire along a
continuous process. LogicComm can summarize each pseudotime bin separately and
plot dynamic cell-type-pair or role trajectories.

**中文说明。** 如果研究对象有分化、激活或时间序列过程，可以把细胞按 pseudotime 或时间
分箱，观察通讯轴和细胞角色是否随状态变化而重排。

```{r pseudotime_dynamics, eval=FALSE}
# pseudotime_vector must be named by cell barcode. It can be numeric or a
# categorical time/stage label.
dyn <- summarize_communication_dynamics(
  reo_mat      = reo_pbmc,
  pseudotime   = pseudotime_vector,
  cell_labels  = celltype_labels,
  lr_db        = lr_pairs_human,
  n_bins       = 6,
  bin_method   = "quantile",
  min_cells_per_bin = 50,
  lcs_threshold = 0.01,
  min_edges     = 20
)

plot_communication_dynamics(dyn, level = "pair", metric = "sum_lcs", top_n = 8)
plot_communication_dynamics(dyn, level = "role", metric = "hub_score", top_n = 8)
```

A trajectory result should be interpreted as descriptive unless the pseudotime is
well validated. The most defensible claims combine a changing L-R axis with a
matching receiver-response program, stable edge support, and external biological
knowledge.


### 9.9 Cell-type communication participation profiles (0.8.0)

Not every cell in a cluster communicates. `summarize_celltype_participation()`
reports, for each cell type, the fraction of cells that express the
ligand/receptor machinery, the pathway and L-R composition of its outgoing and
incoming communication, and whether it is a major communication hub.

**中文说明。** 这一步回答"群里有多少细胞真在通讯、通讯走哪些通路/配受体对、它是不是
重要通讯亚群"。占比按配/受体 REO 激活态计算，不依赖邻域图。

```{r celltype_participation, eval=FALSE}
part <- summarize_celltype_participation(ct_comm, reo_mat = reo_pbmc)
part   # communicating-cell fractions + major-hub label per cell type

# Fraction of cells engaged in communication (sender / receiver) per cell type.
plot_celltype_participation(part)

# Pathway composition of each cell type's outgoing communication.
plot_celltype_pathway_composition(part, direction = "outgoing", level = "pathway")

# One subgroup's profile card: the pathways and L-R pairs it sends and receives.
top_hub <- part$group_participation$cell_type[1]
plot_celltype_communication_profile(part, cell_type = top_hub)
```

### 9.10 Integrated discovery ranking and landscape (0.8.0)

`rank_communication_axes()` (computed as `ranked_axes` above) integrates
communication strength, cell-type-pair specificity, threshold stability
(`sens`), cell-label permutation support (`null_pair`), and receiver response
into one `discovery_score` with an `evidence_tier` (Tier 1-4).
`plot_communication_discovery()` shows the strength-versus-specificity landscape
coloured by tier, so strong, specific, supported axes stand out in the upper
right.

```{r discovery_landscape, eval=FALSE}
head(ranked_axes[, c("sender_type", "receiver_type", "lr_pair",
                     "strength_score", "specificity_score",
                     "discovery_score", "evidence_tier", "broad_axis_flag")], 20)

plot_communication_discovery(ranked_axes)
```

### 9.11 Two confounds that dominate raw rankings, and how to remove them

A raw discovery ranking is dominated by two well-understood confounds. LogicComm
flags both, and acts automatically on the first, so genuinely pair-specific
biology is not buried under them. **Do not rank by raw `lcs`** -- that re-surfaces
exactly these confounds; rank by `discovery_score` (the default order) and use the
flag columns below.

**(a) Broad / ubiquitous and identity axes (handled automatically since 0.9.3).**
`MHC-I -> CD8A`, `CD99 - CD99`, `B2M`, and similar interactions are expressed by
almost every cell, so they earn a high LCS everywhere without being
cell-type-pair specific. `score_communication_specificity()` flags these
(`ubiquitous_interaction_flag`, `specificity_class`). Since 0.9.3,
`rank_communication_axes(demote_broad = TRUE)` (the default) multiplies their
`discovery_score` by `broad_penalty` (default 0.5) and caps their `evidence_tier`
at `Tier 3: broad / non-specific`, so a broad axis can never outrank a specific
candidate. This is ranking-only: the `active` flag and LCS are untouched, and
`demote_broad = FALSE` restores the previous behaviour. The demotion keys on
*ubiquity*, not identity, so a genuine identity-but-specific axis such as
`HLA-E -> KLRC1 / NKG2A` (NK self-recognition) is preserved while `HLA -> CD8A`
is demoted.

**中文说明。** `MHC-I -> CD8A`、`CD99 - CD99`、`B2M` 这类轴几乎所有细胞都表达，LCS
到处都高但没有细胞类型对特异性。0.9.3 起 `rank_communication_axes()` 默认把
`ubiquitous` 轴的 `discovery_score` 乘惩罚、`evidence_tier` 封顶到 Tier 3，所以它们
不会再压过真正特异的候选；但**只针对“广谱”而非“身份”**，因此像 `HLA-E -> NKG2A`
这种真实的 NK 自我识别轴会被保留，而 `HLA -> CD8A` 会被降级。**不要按原始 `lcs` 排序**，
那会把这些混杂重新顶到最前；应按 `discovery_score`（默认顺序）排序。

**(b) The proliferation / transcriptional-breadth hub (diagnostic).** Cycling
cells (high S/G2M, very broad transcriptomes) call many genes logic-active and
form active edges with almost every partner, so a `Cycling_*` cluster can appear
in the large majority of top axes. `diagnose_proliferation_confound()` (run in
Section 9.6) scores S/G2M signatures and transcriptional breadth per cell, flags
breadth/cycling-outlier cell types, and annotates each axis with
`proliferation_confound_flag`. It is diagnostic only -- it does not change any
score -- so you filter it yourself.

```{r confound_clean_view, eval=FALSE}
# Per-cell-type proliferation/breadth summary (which cell types are hubs).
ct_comm$proliferation_summary

# How much do the two confounds dominate the top of the raw ranking?
top50 <- head(ranked_axes, 50)
mean(top50$broad_axis_flag %in% TRUE)              # fraction of top-50 that are broad
mean(top50$proliferation_confound_flag %in% TRUE) # fraction touching a cycling hub

# Recommended one call: drop broad + proliferation-hub axes AND re-scale / re-tier
# the survivors within the kept set, so genuine low-LCS candidates are tiered on
# their own merits instead of against the noise that was removed.
clean_axes <- communication_discovery_view(ranked_axes)   # drop_broad + drop_proliferation
# Shortcut: communication_discovery_view(ct_comm) also works -- it ranks the axes
# on the fly (without permutation support; pass a ranked data.frame for null_support).
head(clean_axes[, c("sender_type", "receiver_type", "lr_pair", "lcs",
                    "discovery_score", "specificity_class", "evidence_tier")], 20)

# To also drop canonical identity-recognition axes (e.g. HLA-E -> NKG2A) and keep
# only novel candidates:
novel_axes <- communication_discovery_view(ranked_axes, drop_identity = TRUE)

# rescale = FALSE is the plain filter equivalent (keeps the original scores/tiers):
clean_manual <- subset(ranked_axes,
                       !broad_axis_flag %in% TRUE &
                       !proliferation_confound_flag %in% TRUE)
```

**中文说明。** 增殖/转录广度中枢（如 `Cycling_*` 簇）会出现在绝大多数顶端轴里，但这
反映的是转录广度而非特异通讯。`communication_discovery_view()` 一步完成推荐的发表级
发现视图：去掉 `broad_axis_flag` 与 `proliferation_confound_flag`，并在**保留集合内部
重新缩放 strength/specificity、重判 evidence_tier**。这是必要的——主排名对全表缩放，
广谱/增殖这些高 LCS 轴把"天花板"抬高，导致真正特异但 LCS 偏低的候选只能落到 Tier 3/4；
在保留集合内重判后，它们才会按自身证据获得合理的 Tier 1/2。加 `drop_identity = TRUE`
会进一步去掉 `HLA-E -> NKG2A` 这类经典识别轴，只留下新颖候选（如 `CCL5 -> CCR4`、
`TNFSF9 -> TNFRSF9` 共刺激、`MIF -> CD74`、`ALCAM-CD6` 等）。`rescale = FALSE` 时则
等价于上面的手动 `subset()`，只过滤、不重判。

> **Caveat -- cell order.** `diagnose_proliferation_confound()` matches `expr`
> columns to `ct_comm$cell_labels` by cell name when both are named; otherwise it
> requires the same number of cells in the same order. Use the same cells you used
> to build `reo_mat` / `ct_comm`. The default S/G2M sets are the Tirosh et al.
> (2016) lists (override via `s_genes` / `g2m_genes`); genes absent from `expr`
> are dropped, and z-scored hub flags need several cell types to be meaningful.

### 9.12 Publication figures: styling, controls, and export

Every `plot_*` function shares one colorblind-safe, print-safe LogicComm theme
and brand palette, so the whole figure suite is consistent by default. You can
restyle custom layers, tune the key figures, and export at journal widths.

**Brand scales for custom ggplot layers:**

- `theme_logiccomm(grid = "xy"/"x"/"y"/"none")` — the shared theme
- `scale_color/fill_logiccomm_c()` — sequential (brand "mako" ramp)
- `scale_color/fill_logiccomm_diverging(midpoint = 0)` — diverging (teal–amber)
- `scale_color/fill_logiccomm_d()` — qualitative (categorical)
- `scale_color/fill_tier()` — the ordered `evidence_tier` scale
- `logiccomm_brand$primary/secondary/accent/highlight/ink` — brand hex colours

**Key per-figure publication controls** (added in the 0.10.x figure overhaul):

| Figure | Useful arguments |
|---|---|
| `plot_communication_discovery()` | `title = NULL`, `subtitle`, `top_n_label` |
| `plot_differential_celltype_volcano()` | `fdr_cutoff`, `lfc_threshold`, `x_lab` |
| `plot_celltype_glm_volcano()` | `fdr_cutoff`, `effect_threshold` |
| `plot_celltype_network()` / `plot_celltype_network_publication()` | `layout = "auto"/"fr"/"circle"` |
| `plot_lr_bubble_advanced()` | `top_n_pathways` (caps the pathway facets) |
| `plot_pathway_heatmap()` | `cluster_rows`, `cluster_cols` (default `TRUE`) |
| `plot_celltype_roles()`, `plot_communication_discovery()` | `subtitle` |

```{r figure_publication, eval=FALSE}
# Tune key figures
vol <- plot_differential_celltype_volcano(ct_comparison, fdr_cutoff = 0.05, lfc_threshold = 1)
net <- plot_celltype_network_publication(ct_comm, layout = "fr")
bub <- plot_lr_bubble_advanced(ct_comm, top_n_pathways = 9)

# Drop the in-plot title for a caption-driven journal layout
disc <- plot_communication_discovery(ranked_axes, title = NULL)

# Export at journal column widths. Vector PDF/SVG is recommended; PDF uses
# cairo_pdf so the arrow/ellipsis glyphs in axis labels embed cleanly.
save_logiccomm_figure(disc, "discovery.pdf", width = "single")                 # 89 mm
save_logiccomm_figure(net,  "network.png",  width = "double", height_mm = 120, dpi = 600)

# Multi-panel figure (a/b/c) with patchwork
library(patchwork)
fig <- plot_communication_discovery(ranked_axes) + net +
       patchwork::plot_annotation(tag_levels = "a")
save_logiccomm_figure(fig, "figure2.pdf", width = "double", height_mm = 120)
```

**中文说明。** 所有 `plot_*` 默认套用统一的色盲安全品牌主题与配色，整套图风格一致。
自定义图层可复用上面的品牌色阶与 `logiccomm_brand` 颜色；关键图各自暴露了发表常用参数
（volcano 的 `fdr_cutoff` / `lfc_threshold`、network 的 `layout`、bubble 的
`top_n_pathways`、discovery/roles 的 `subtitle`、heatmap 的聚类等）。用
`save_logiccomm_figure()` 按期刊栏宽（单栏 89 mm / 双栏 183 mm）导出矢量图（PDF 默认走
cairo，箭头/省略号等字形可正确嵌入），用 `patchwork` 拼多面板（a/b/c）；`title = NULL`
去掉图内标题改用图注。

## Step 10: Optional Cell-Type-Level Case vs Control Comparison

For multi-sample studies, compute cell-type-resolved communication inside each
biological sample, then compare those sample-level vectors. Do not pool all cells
across samples as the statistical unit.

**中文说明。** 多样本研究要把“样本”作为统计单位。正确做法是每个样本单独计算
cell-type-resolved communication，再比较样本级向量；不要把所有细胞合并后直接做
Case vs Control，否则会产生伪重复。

```{r multisample_celltype_comm, eval=FALSE}
ct_lcs_list <- lapply(names(pbmc_list), function(sname) {
  reo_i <- calc_REO_matrix(
    pbmc_list[[sname]],
    lr_genes = lr_genes,
    layer = "counts",
    verbose = FALSE
  )

  labels_i <- setNames(
    as.character(pbmc_list[[sname]][[label_col]][, 1]),
    colnames(pbmc_list[[sname]])
  )

  ct_i <- summarize_celltype_communication(
    reo_i,
    cell_labels      = labels_i,
    lr_db            = lr_pairs_human,
    lcs_threshold    = 0.01,
    min_edges        = 20,
    verbose          = FALSE
  )

  celltype_comm_to_lcs(ct_i, level = "celltype_lr", metric = "lcs")
})
names(ct_lcs_list) <- names(pbmc_list)

ct_comparison <- CompareLogicGroups(
  ct_lcs_list,
  group_info    = group_info,
  case_label    = "Case",
  ctrl_label    = "Ctrl",
  lcs_threshold = 0.01,
  verbose       = TRUE
)

head(ct_comparison[order(ct_comparison$fdr_fisher), ], 20)

plot_differential_celltype_heatmap(ct_comparison, metric = "asymmetry")
plot_differential_celltype_volcano(ct_comparison, x = "log2fc_lcs")
```

This table asks whether specific celltypeA -> celltypeB -> L-R features are
recurrently active in Case samples relative to Control samples. For example, a
Case-enriched `CD14 Mono|CD4 T|CXCL12_CXCR4` feature would support a
sample-recurrent myeloid-to-T-cell communication hypothesis.

### 10.1 Sample-level GLM with covariates

For publication-level cohort studies, a simple frequency comparison is often a
useful first pass, but a sample-level model is preferable when covariates or
unequal edge opportunities matter. `fit_celltype_comm_glm()` models active edge
counts using available edge opportunities as trials.

**中文说明。** 如果样本量足够并且有批次、年龄、性别等协变量，GLM 比简单频率比较更适合
发表级队列分析。它把 active edge count 作为事件数，把可用 edge opportunities 作为试验数。

```{r multisample_glm, eval=FALSE}
# sample_ct_list should contain the full LogicCommCellTypeComm object per sample,
# not only the vector returned by celltype_comm_to_lcs().
sample_meta <- data.frame(
  group = group_info[names(sample_ct_list)],
  batch = batch_info[names(sample_ct_list)],
  row.names = names(sample_ct_list)
)

glm_res <- fit_celltype_comm_glm(
  sample_ct_list,
  sample_metadata = sample_meta,
  design = ~ group + batch,
  coef = "groupCase",
  level = "celltype_lr",
  min_samples = 4,
  min_total_trials = 50
)

head(glm_res[order(glm_res$fdr), ], 20)
plot_celltype_glm_volcano(glm_res)
```

The GLM coefficient is a log-odds effect for active edge support. A positive
`groupCase` estimate means the feature has higher active-edge probability in
Case after adjusting for the supplied covariates.

---

### 10.2 Subgroup-resolved differential communication (0.8.0)

`differential_celltype_communication()` reorganizes the cohort comparison by the
sender -> receiver subgroup on which each L-R change occurs, so the cell-type
pairs that carry the differential signaling are directly visible.

**中文说明。** 这一步直接回答"主要差异通讯对发生在哪些亚群上"。注意小样本时
Fisher FDR 会饱和到 1，应按效应量（asymmetry / log2fc）排序，并配合
`fit_celltype_comm_glm()` 与更多生物学重复，而不要把 FDR 当作"无差异"的证据。

```{r differential_subgroup, eval=FALSE}
# sample_ct_list: named list of per-sample LogicCommCellTypeComm objects (Step 10).
diff <- differential_celltype_communication(
  sample_ct_list, group_info,
  case_label = "Case", ctrl_label = "Ctrl"
)
diff                     # top differential sender -> receiver subgroups + FDR caveat

# How many L-R pairs change per sender -> receiver subgroup, split by direction.
plot_differential_communication_summary(diff)

# sender -> receiver subgroup x L-R pair, coloured by Case/Ctrl asymmetry.
plot_differential_celltype_heatmap(diff, metric = "asymmetry")
```

## Step 11: Optional Cell Type Annotation

PBMC3K is often annotated using canonical markers. The following is only a
minimal example; adjust labels after inspecting marker genes.

**中文说明。** 细胞类型注释会直接影响 sender/receiver 的解释。这里给的是最小示例，
真实分析中应结合 marker gene、已有注释和生物学背景认真校正。

```{r marker_plot, eval=FALSE, fig.width=9, fig.height=6}
FeaturePlot(
  pbmc,
  features = c("MS4A1", "CD3D", "LYZ", "NKG7", "PPBP", "FCGR3A"),
  reduction = "umap",
  ncol = 3
)
```

```{r manual_labels, eval=FALSE}
# Example only: replace these mappings after inspecting your clustering result.
# names(new_ids) must match current cluster IDs.
new_ids <- c(
  "0" = "CD4 T / mixed T",
  "1" = "CD14 Mono",
  "2" = "B cell",
  "3" = "CD8 T / NK",
  "4" = "FCGR3A Mono",
  "5" = "Dendritic",
  "6" = "Megakaryocyte"
)

pbmc$cell_type_demo <- plyr::mapvalues(
  x    = as.character(pbmc$seurat_clusters),
  from = names(new_ids),
  to   = unname(new_ids),
  warn_missing = FALSE
)

DimPlot(pbmc, group.by = "cell_type_demo", label = TRUE) + NoLegend()
VlnPlot(pbmc, features = "LogicComm_score", group.by = "cell_type_demo", pt.size = 0)
```

**Biological interpretation.** Canonical PBMC markers make the communication
results easier to explain. `MS4A1` marks B cells, `CD3D` marks T cells, `LYZ`
marks monocytes/dendritic-like myeloid cells, `NKG7` marks NK/cytotoxic cells,
`PPBP` marks platelet/megakaryocyte-like cells, and `FCGR3A` marks non-classical
monocytes. After assigning cell-type labels, a high LogicComm score in a named
cell type can be interpreted as a cell-type-level communication program rather
than an anonymous Seurat cluster effect.

If you use the optional manual annotation block, install `plyr` first:

```{r install_plyr, eval=FALSE}
install.packages("plyr")
```

---

## Step 12: Manual PBMC3K Download Alternative

If `SeuratData` is unavailable, download PBMC3K directly from 10X Genomics:

**中文说明。** 如果 `SeuratData` 安装失败，可以直接下载 10X 官方 PBMC3K 数据，手动创建
Seurat 对象，然后从 Step 2 继续。

```{r manual_download, eval=FALSE}
dir.create("data", showWarnings = FALSE)

url <- "https://cf.10xgenomics.com/samples/cell-exp/1.1.0/pbmc3k/pbmc3k_filtered_gene_bc_matrices.tar.gz"
dest <- file.path("data", "pbmc3k_filtered_gene_bc_matrices.tar.gz")
download.file(url, destfile = dest, mode = "wb")
untar(dest, exdir = "data")

pbmc_data <- Read10X(data.dir = "data/filtered_gene_bc_matrices/hg19")
pbmc <- CreateSeuratObject(counts = pbmc_data, project = "pbmc3k")
```

Then continue from **Step 2: Standard Seurat Preprocessing**.

---

## Step 13: Multiple PBMC Samples

In real projects you usually have multiple independent PBMC samples, for example:

**中文说明。** 真实项目通常有多个 PBMC 样本。LogicComm 推荐每个样本独立预处理、独立计算
REO/LCS，然后在样本层面比较，而不是把细胞混成一个“大样本”。

```text
pbmcsample001, pbmcsample002, pbmcsample003, ...
```

LogicComm treats each biological sample as one independent unit. The recommended
workflow is:

1. Preprocess each Seurat object separately.
2. Run `calc_REO_matrix()` and `IdentifyLogicConsensus()` per sample.
3. Compare LCS across groups with `CompareLogicGroups()` or `run_multisample()`.

Do **not** merge all samples into one Seurat object before LogicComm group
comparison if your goal is Case vs Control sample-level inference. Merging cells
can turn biological replicates into pseudo-replicates. Keep one Seurat object per
sample.

### 13.1 Prepare a named list of Seurat objects

Suppose you already have several Seurat objects in R:

```{r multisample_objects, eval=FALSE}
pbmc_list <- list(
  pbmcsample001 = pbmcsample001,
  pbmcsample002 = pbmcsample002,
  pbmcsample003 = pbmcsample003,
  pbmcsample004 = pbmcsample004,
  pbmcsample005 = pbmcsample005,
  pbmcsample006 = pbmcsample006
)
```

Define the sample group information. Names must match `names(pbmc_list)`:

```{r multisample_groups, eval=FALSE}
group_info <- c(
  pbmcsample001 = "Case",
  pbmcsample002 = "Case",
  pbmcsample003 = "Case",
  pbmcsample004 = "Ctrl",
  pbmcsample005 = "Ctrl",
  pbmcsample006 = "Ctrl"
)
```

Check the alignment before running:

```{r multisample_check_names, eval=FALSE}
stopifnot(identical(names(pbmc_list), names(group_info)))
table(group_info)
```

### 13.2 Preprocess each sample separately

Each Seurat object should have its own KNN graph. A small helper function makes
this easier:

```{r multisample_preprocess_fun, eval=FALSE}
preprocess_pbmc_for_logiccomm <- function(obj, dims = 1:20, resolution = 0.5) {
  obj <- NormalizeData(obj, verbose = FALSE)
  obj <- FindVariableFeatures(obj, selection.method = "vst", nfeatures = 2000,
                              verbose = FALSE)
  obj <- ScaleData(obj, verbose = FALSE)
  obj <- RunPCA(obj, npcs = max(dims), verbose = FALSE)
  obj <- FindNeighbors(obj, dims = dims, verbose = FALSE)
  obj <- RunUMAP(obj, dims = dims, verbose = FALSE)
  obj <- FindClusters(obj, resolution = resolution, verbose = FALSE)
  obj
}

pbmc_list <- lapply(pbmc_list, preprocess_pbmc_for_logiccomm)
```

Confirm that every sample has a nearest-neighbor graph:

```{r multisample_graph_check, eval=FALSE}
lapply(pbmc_list, function(x) names(x@graphs))
```

For standard RNA workflows, you will usually see `RNA_nn` and `RNA_snn`. For
LogicComm KNN mode, `RNA_nn` is recommended because it represents the direct
nearest-neighbor adjacency.

### 13.3 Recommended explicit multi-sample workflow

`run_multisample()` processes a list of Seurat objects and computes each sample's
LCS from REO co-expression (no per-cell graph). The explicit workflow below is
still shown because it is transparent and makes it easy to save each sample's REO
matrix, LCS vector, and optional cell-type communication summary:

```{r multisample_explicit, eval=FALSE}
data(lr_pairs_human)
lr_genes <- all_lr_genes(lr_pairs_human)

reo_list <- lapply(names(pbmc_list), function(sname) {
  message("[", sname, "] REO")
  calc_REO_matrix(
    pbmc_list[[sname]],
    lr_genes       = lr_genes,
    rank_threshold = 0.5,
    layer          = "counts",
    chunk_size     = 2000,
    verbose        = FALSE
  )
})
names(reo_list) <- names(pbmc_list)

lcs_list <- lapply(names(pbmc_list), function(sname) {
  message("[", sname, "] LCS")
  IdentifyLogicConsensus(
    reo_list[[sname]],
    lr_db         = lr_pairs_human,
    lcs_threshold = 0.01,
    verbose       = FALSE
  )
})
names(lcs_list) <- names(pbmc_list)
```

Now compare Case vs Ctrl at the sample level:

```{r multisample_compare, eval=FALSE}
comparison <- CompareLogicGroups(
  lcs_list,
  group_info    = group_info,
  case_label    = "Case",
  ctrl_label    = "Ctrl",
  lcs_threshold = 0.01,
  verbose       = TRUE
)

head(comparison, 20)
```

**Biological interpretation.** In a real Case-vs-Ctrl PBMC study, this table asks
which LR programs are recurrently active in one group of biological samples but
not the other. A Case-enriched cytokine or chemokine pair may suggest disease- or
treatment-associated immune activation. A Ctrl-enriched pair may indicate a
resting/homeostatic program lost in disease. Because the unit of comparison is
the sample, not the individual cell, the result is more appropriate for cohort
inference than pooling all cells together.

Visualize the group comparison:

```{r multisample_plots, eval=FALSE, fig.width=12, fig.height=5}
plot_lcs_bubble(
  comparison,
  top_n        = 20,
  min_asymmetry = 0.2,
  color_by     = "asymmetry"
)

plot_lcs_heatmap(
  comparison,
  group_info = group_info,
  top_n      = 30,
  fdr_cutoff = 0.5
)
```

**Biological interpretation.** The bubble plot highlights LR pairs with strong
case-control asymmetry; the heatmap shows whether those signals are consistent
across individual samples. A convincing biological signal should usually show
both meaningful group asymmetry and reasonable within-group recurrence. If one
sample dominates the heatmap, interpret the result as sample-specific until more
replicates or orthogonal validation support it.

Export results:

```{r multisample_export, eval=FALSE}
export_results(comparison, file = "PBMC_multi_sample_LogicComm_comparison.csv")
```

### 13.4 One-line helper workflow

`run_multisample()` runs per-sample REO + co-expression LCS + group comparison
in one call:

```{r multisample_run_helper, eval=FALSE}
result <- run_multisample(
  pbmc_list,
  group_info     = group_info,
  lr_db          = lr_pairs_human,
  rank_threshold = 0.5,
  lcs_threshold  = 0.01,
  case_label     = "Case",
  ctrl_label     = "Ctrl",
  layer          = "counts",
  chunk_size     = 2000,
  mc_cores       = 1,
  verbose        = TRUE
)

result
comparison <- result$comparison
```

On Windows, keep `mc_cores = 1`. On macOS/Linux, you may use `mc_cores > 1` for
parallel processing if memory is sufficient.

### 13.5 If you already have matrices instead of Seurat objects

You can also analyze a named list of raw count matrices. Each matrix must be
**genes x cells**:

```{r multisample_matrix, eval=FALSE}
expr_list <- list(
  pbmcsample001 = counts001,
  pbmcsample002 = counts002,
  pbmcsample003 = counts003,
  pbmcsample004 = counts004
)

result <- run_multisample(
  expr_list,
  group_info     = group_info[names(expr_list)],
  lr_db          = lr_pairs_human,
  rank_threshold = 0.5,
  lcs_threshold  = 0.01,
  mc_cores       = 1
)
```

Each sample's LCS is computed from REO co-expression; any legacy `knn_list` /
graph arguments are deprecated and ignored.

### 13.6 Practical recommendations for multiple PBMC samples

**中文说明。** 多样本分析的核心原则是：样本是统计重复，细胞不是独立病人。所有样本应使用
相同 LR 数据库、相同 REO 阈值、相同 assay/layer，并保持样本名与分组信息严格对齐。

- Use one object per biological sample.
- Keep sample names unique and stable: `pbmcsample001`, `pbmcsample002`, etc.
- Keep `group_info` named and aligned to `pbmc_list`.
- Use the same `rank_threshold`, LR database, assay, and layer for all samples.
- Do not compare only cell-pooled results; use sample-level replicates.
- For Windows, use `mc_cores = 1`.
- For large samples, reduce `chunk_size`, for example `chunk_size = 1000`.

---

## Complete Minimal Script

**中文说明。** 如果你只想快速跑通流程，可以先运行下面这个最小脚本；前面章节提供的是
每一步的解释和扩展分析。

```{r minimal_script, eval=FALSE}
library(LogicComm)
library(Seurat)
library(SeuratData)

SeuratData::data("pbmc3k")
pbmc <- pbmc3k

pbmc <- NormalizeData(pbmc, verbose = FALSE)
pbmc <- FindVariableFeatures(pbmc, verbose = FALSE)
pbmc <- ScaleData(pbmc, verbose = FALSE)
pbmc <- RunPCA(pbmc, npcs = 30, verbose = FALSE)
pbmc <- FindNeighbors(pbmc, dims = 1:20, verbose = FALSE)
pbmc <- RunUMAP(pbmc, dims = 1:20, verbose = FALSE)
pbmc <- FindClusters(pbmc, resolution = 0.5, verbose = FALSE)

data(lr_pairs_human)
lr_genes <- all_lr_genes(lr_pairs_human)

reo_pbmc <- calc_REO_matrix(
  pbmc,
  lr_genes       = lr_genes,
  rank_threshold = 0.5,
  layer          = "counts"
)

lcs_pbmc <- IdentifyLogicConsensus(reo_pbmc, lr_db = lr_pairs_human)

head(sort(lcs_pbmc[!is.na(lcs_pbmc)], decreasing = TRUE), 20)

cell_scores <- score_lr_activity(
  reo_pbmc,
  lr_db     = lr_pairs_human,
  mode      = "both",
  aggregate = TRUE
)

cell_order <- colnames(pbmc)
pbmc$LogicComm_score <- cell_scores$comm_score[cell_order]

FeaturePlot(pbmc, features = "LogicComm_score", reduction = "umap")
```

---

## Troubleshooting

**中文说明。** 这一节列出常见问题：PBMC3K 下载失败、KNN 图找不到、细胞名不匹配、内存过高、
以及应该使用哪个 assay layer。

### `SeuratData::InstallData("pbmc3k")` fails

Use the manual 10X download route in Step 12.

### Warning: a `knn_mat` / `mode` / `graph_*` argument is deprecated and ignored

v0.12 scores at the cell-type level from co-expression and no longer uses a
per-cell KNN/SNN graph. Drop these arguments — they are accepted only for
backward compatibility and have no effect. (Cell-type labels still come from your
usual Seurat clustering; LogicComm just doesn't consume the graph.)

### Cell names do not match between REO and KNN

Make sure `reo_pbmc` and the Seurat graph are generated from the same Seurat
object without subsetting one but not the other.

```{r troubleshoot_names, eval=FALSE}
all(colnames(reo_pbmc) %in% colnames(pbmc))
length(colnames(reo_pbmc))
ncol(pbmc)
```

### Memory usage is high

Use a smaller chunk size:

```{r troubleshoot_memory, eval=FALSE}
reo_pbmc <- calc_REO_matrix(
  pbmc,
  lr_genes   = lr_genes,
  chunk_size = 1000
)
```

### Which layer should I use?

For LogicComm's binary REO logic, raw counts are usually a good default:

```{r layer_counts, eval=FALSE}
reo_pbmc <- calc_REO_matrix(pbmc, lr_genes = lr_genes, layer = "counts")
```

You may also try normalized data if your study design requires it, but keep the
same choice across all samples.

---

## Session Info

```{r sessioninfo, eval=FALSE}
sessionInfo()
```
