---
title: "LogicComm: multi-sample cell-type co-expression analysis"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Overview

**LogicComm** converts single-cell expression into Relative Expression Ordering
(REO) logic states and scores ligand-receptor communication with Logic Consensus
Scores (LCS). The recommended public API is the `logic_*` pipeline:

1. `logic_check_lrdb()` / `logic_import_lrdb()` validate or import an LR database.
2. `logic_prepare()` builds REO logic matrices.
3. `logic_score_lr()` scores sample-level LR logic.
4. `logic_summarize_celltypes()` summarizes sender-cell-type -> receiver-cell-type
   communication.
5. `logic_run()` and `logic_compare_groups()` compare biological samples.

LogicComm 0.12 centers ordinary dissociated scRNA-seq analysis on **cell-type
co-expression**, not on per-cell KNN/SNN neighborhood aggregation. A Seurat graph is
useful for clustering and UMAP, but it is not physical proximity evidence. For
cell-type interpretation, LogicComm asks whether a ligand is active in sender-type
cells and a receptor is active in receiver-type cells.

**中文说明。** LogicComm 的核心流程是 `logic_*`。普通 scRNA-seq 中，核心证据是
cell-type co-expression，而不是把 Seurat KNN/SNN 当作空间邻域。旧函数仍保留给历史脚本，
但新文档优先使用 `logic_prepare()`、`logic_score_lr()`、`logic_summarize_celltypes()`、
`logic_run()`。

---

## Installation

```{r install, eval=FALSE}
install.packages(c(
  "Matrix", "dplyr", "ggplot2", "ggrepel", "scales", "Rcpp"
))

# Optional plotting / Seurat helpers
install.packages(c("pheatmap", "Seurat", "SeuratObject", "patchwork"))

install.packages("LogicComm_0.12.0.tar.gz", repos = NULL, type = "source")
```

Input matrices should be **genes x cells**. Sparse matrices are supported. On
Windows, keep `mc_cores = 1` unless you provide another parallel backend.

---

## Simulated data

This tutorial simulates 8 samples (4 Case + 4 Control). Case samples have elevated
VEGFA and IL6 axes so the expected direction is known.

```{r simulate}
library(LogicComm)
set.seed(2025)

n_cells <- 300
n_genes <- 80
gene_pool <- c(
  "VEGFA", "KDR", "FLT1", "TGFB1", "TGFBR1", "TGFBR2",
  "CXCL12", "CXCR4", "TNF", "TNFRSF1A", "IL6", "IL6R", "IL6ST",
  "COL1A1", "CD44", "PDGFB", "PDGFRB", "FGF2", "FGFR1",
  "CD274", "PDCD1", "HGF", "MET", "ANGPT1", "TEK",
  "CCL2", "CCR2", "CXCL10", "CXCR3",
  paste0("BG", seq_len(n_genes - 28))
)

make_sample <- function(seed, case = FALSE) {
  set.seed(seed)
  mat <- matrix(
    stats::rpois(length(gene_pool) * n_cells, lambda = 1.2),
    nrow = length(gene_pool),
    ncol = n_cells,
    dimnames = list(gene_pool, paste0("C", seed, "_", seq_len(n_cells)))
  )
  if (case) {
    mat["VEGFA", ] <- mat["VEGFA", ] + stats::rpois(n_cells, 6)
    mat["KDR", ]   <- mat["KDR", ]   + stats::rpois(n_cells, 4)
    mat["IL6", ]   <- mat["IL6", ]   + stats::rpois(n_cells, 5)
    mat["IL6R", ]  <- mat["IL6R", ]  + stats::rpois(n_cells, 3)
  }
  mat
}

sample_list <- c(
  lapply(1:4, function(s) make_sample(s, case = TRUE)),
  lapply(5:8, function(s) make_sample(s + 10, case = FALSE))
)
names(sample_list) <- c(paste0("Case_", 1:4), paste0("Ctrl_", 1:4))

group_info <- c(rep("Case", 4), rep("Ctrl", 4))
names(group_info) <- names(sample_list)
```

---

## Step 1: Validate the LR database

```{r lrdb}
data(lr_pairs_human)
logic_check_lrdb(lr_pairs_human, require_metadata = TRUE)
lr_genes <- logic_get_lr_genes(lr_pairs_human)
head(lr_genes)
```

`lr_pairs_human` is a compact LR seed database. If you use a custom database, keep
its schema stable across all samples and validate it before scoring.

---

## Step 2: Build one REO matrix

```{r reo}
reo_demo <- logic_prepare(
  sample_list[[1]],
  lr_db          = lr_pairs_human,
  rank_threshold = 0.5,
  gene_background = "quantile",
  return_rank    = TRUE,
  verbose        = FALSE
)
reo_demo
```

`logic_prepare()` computes a within-cell REO anchor and returns LR genes for
scoring. With `return_rank = TRUE`, downstream cell-type summaries can use
rank-weighted LCS.

---

## Step 3: Score sample-level LCS

```{r lcs}
lcs_demo <- logic_score_lr(
  reo_demo,
  lr_db         = lr_pairs_human,
  lcs_threshold = 0.01,
  verbose       = FALSE
)

sort(lcs_demo, decreasing = TRUE)[1:10]
```

The result is a named vector: one LCS value per LR pair. This sample-level vector
is useful for multi-sample comparison but does not yet assign sender and receiver
cell types.

---

## Step 4: Multi-sample pipeline

`logic_run()` wraps per-sample REO building, per-sample LCS scoring, and group
comparison.

```{r multisample, message=TRUE}
result <- logic_run(
  sample_list,
  group_info     = group_info,
  lr_db          = lr_pairs_human,
  rank_threshold = 0.5,
  lcs_threshold  = 0.01,
  chunk_size     = 1000,
  mc_cores       = 1,
  verbose        = TRUE
)
```

```{r print_result}
print(result$comparison, n = 10)
```

You can retrieve or recompute the comparison with `logic_compare_groups()`:

```{r compare_groups}
comparison <- logic_compare_groups(result)
head(comparison[order(-abs(comparison$asymmetry)), ], 10)
```

---

## Step 5: Visualize group-level results

```{r bubble, fig.cap="Case vs Control frequency bubble plot. Points above the diagonal are enriched in Case. Bubble size = |asymmetry|."}
plot_lcs_bubble(result$comparison, top_n = 15, min_asymmetry = 0.2)
```

```{r heatmap, fig.height=6, eval=requireNamespace("pheatmap", quietly = TRUE), fig.cap="LCS values across samples for top differentially active pairs."}
plot_lcs_heatmap(
  result$comparison,
  group_info = group_info,
  top_n      = 20,
  fdr_cutoff = 1
)
```

This simulation is underpowered for publication-level FDR. Use it to check
direction, recurrence, and effect size.

---

## Step 6: Cell-type co-expression summary

If cell labels are available, summarize sender-cell-type -> receiver-cell-type
communication. This example uses mock labels only to demonstrate the API.

```{r celltype_comm_intro, eval=FALSE}
cell_labels <- setNames(
  rep(c("Fibroblast", "Endothelial", "Macrophage"), length.out = n_cells),
  colnames(sample_list[[1]])
)

ct_comm <- logic_summarize_celltypes(
  reo_demo,
  cell_labels      = cell_labels,
  lr_db            = lr_pairs_human,
  include_self     = TRUE,
  lcs_threshold    = 0.01,
  min_edges        = 20,
  min_expr_frac    = 0.1,
  lcs_weighting    = "rank",
  verbose          = TRUE
)

head(ct_comm$pair_summary[order(-ct_comm$pair_summary$sum_lcs), ], 10)
head(ct_comm$role_summary, 10)
```

Plot the cell-type result:

```{r celltype_plots_intro, eval=FALSE}
plot_celltype_heatmap(ct_comm, metric = "sum_lcs")
plot_celltype_roles(ct_comm)
plot_celltype_role_heatmap(ct_comm)
plot_lr_bubble_advanced(ct_comm, top_n_per_pair = 5)
```

`include_self = TRUE` keeps same-cell-type communication such as `Macrophage ->
Macrophage`, interpreted as homotypic/autocrine-like potential.

---

## Step 7: Compare cell-type-resolved features across samples

For cohort analysis, compute `ct_comm` inside each biological sample, convert each
object to a named vector, then compare vectors with `logic_compare_groups()`.

```{r celltype_compare_intro, eval=FALSE}
sample_ct_list <- lapply(names(sample_list), function(sname) {
  reo_i <- logic_prepare(
    sample_list[[sname]],
    lr_db          = lr_pairs_human,
    rank_threshold = 0.5,
    gene_background = "quantile",
    return_rank    = TRUE,
    verbose        = FALSE
  )
  labels_i <- setNames(
    rep(c("Fibroblast", "Endothelial", "Macrophage"), length.out = ncol(sample_list[[sname]])),
    colnames(sample_list[[sname]])
  )
  logic_summarize_celltypes(
    reo_i,
    cell_labels   = labels_i,
    lr_db         = lr_pairs_human,
    lcs_threshold = 0.01,
    min_edges     = 20,
    min_expr_frac = 0.1,
    lcs_weighting = "rank",
    verbose       = FALSE
  )
})
names(sample_ct_list) <- names(sample_list)

ct_lcs_list <- lapply(sample_ct_list, celltype_comm_to_lcs,
                      level = "celltype_lr", metric = "lcs")

ct_comparison <- logic_compare_groups(
  ct_lcs_list,
  group_info    = group_info[names(ct_lcs_list)],
  case_label    = "Case",
  ctrl_label    = "Ctrl",
  lcs_threshold = 0.01
)
```

The comparison unit remains the biological sample, not the cell.

---

## Step 8: Per-cell activity scores

```{r cell_score}
reo_logic <- reo_demo$logic
scores <- score_lr_activity(reo_logic, mode = "both", aggregate = TRUE, verbose = FALSE)
summary(scores$comm_score)
```

```{r cell_score_hist}
hist(
  scores$comm_score,
  breaks = 30,
  col = "#64B5F6",
  main = "Per-cell communication score distribution",
  xlab = "Fraction of LR pairs active",
  border = "white"
)
```

```{r rank_cells}
top_cells <- rank_comm_cells(scores, n = 10)
print(top_cells[, c("cell", "sender_score", "receiver_score", "comm_score", "rank")])
```

---

## Step 9: Rank shift analysis

Rank shift asks which genes became more dominant in Case cells, independently of
LR-pair scoring.

```{r rank_shift}
rs <- calc_rank_shift(
  sample_list,
  group_info = group_info,
  genes      = lr_genes
)
print(rs, n = 8)
```

```{r rank_shift_plot, fig.cap="Rank-shift volcano plot."}
plot_rank_shift(rs, top_n_label = 15, min_shift_label = 0.05)
```

---

## Step 10: Filter and export

```{r filter_export, eval=FALSE}
sig_pairs <- filter_lcs(
  result$comparison,
  min_asymmetry = 0.3,
  max_fdr       = 0.05,
  direction     = "up"
)

export_results(sig_pairs, file = "LogicComm_significant_pairs.csv")
```

---

## Interpretation guide

| Output column | Meaning |
|---|---|
| `asymmetry` | `case_freq - ctrl_freq`; positive values are Case-enriched. |
| `case_freq` / `ctrl_freq` | Fraction of samples active above the LCS threshold. |
| `lcs` | Logic Consensus Score for an LR pair or cell-type LR feature. |
| `log2fc_lcs` | Fold-change in mean LCS. Use with recurrence and sample count. |
| `fdr_fisher` | BH-adjusted Fisher p-value on sample-level recurrence. |
| `discovery_score` | Integrated strength/specificity/support score for cell-type axes. |

---

## Current behavior notes

- Multimer ligands/receptors require all subunits to be present and active.
- `group_info` must contain one label for every sample.
- L-R pairs are aligned by name across samples; missing pairs are treated as `NA`.
- `filter_lcs(direction = ...)` accepts only `"up"`, `"down"`, or `"both"`.
- `plot_lcs_bubble(color_by = ...)` accepts `"log2fc"` or `"asymmetry"`.
- Legacy neighborhood/KNN arguments are compatibility inputs; new ordinary
  scRNA-seq analyses should use the cell-type co-expression workflow above.

---

## Troubleshooting

### Missing dependencies during install

```{r install_missing, eval=FALSE}
install.packages(c("dplyr", "ggplot2", "ggrepel", "scales", "Rcpp"))
```

### Cell labels do not align

```{r check_label_names, eval=FALSE}
all(colnames(reo_demo$logic) %in% names(cell_labels))
setdiff(colnames(reo_demo$logic), names(cell_labels))[1:5]
```

### Heatmap dependency

```{r install_pheatmap, eval=FALSE}
install.packages("pheatmap")
```

---

## Session info

```{r sessioninfo}
sessionInfo()
```
