---
title: "PBMC multi-sample simulation demo"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  message = FALSE,
  warning = FALSE
)
```

## Purpose

This tutorial demonstrates the LogicComm 0.12 sample-level workflow using the same
PBMC3K data source as `Seurat_demo.Rmd`. It is a **software validation demo**, not a
biological disease analysis:

1. Split one PBMC3K count matrix into four pseudo-samples: `pbmc1` to `pbmc4`.
2. Treat `pbmc1` and `pbmc2` as simulated Case samples.
3. Treat `pbmc3` and `pbmc4` as simulated Ctrl samples.
4. Add a controlled perturbation to selected LR genes in Case only.
5. Run the recommended `logic_run()` / `logic_compare_groups()` sample-level API.

**中文说明。** 这个示例验证的是软件流程：样本名是否对齐、已知扰动是否能被恢复、比较是否
以“样本”为统计单位。它不能替代真实 Case/Ctrl 队列，也不应把 PBMC3K 的模拟扰动解释成
真实疾病机制。

Because there are only two simulated samples per group, Fisher-test p-values have
very low resolution. Success is judged by direction, recurrence, and effect size:
targeted LR pairs should be active in both Case pseudo-samples, inactive or weaker
in Controls, and have positive `log2fc_lcs`.

---

## Executable script

The full executable script is installed with the package at
`inst/examples/pbmc_multisample_demo.R`. After installing LogicComm and PBMC3K via
SeuratData, run:

```{r run_script, eval=FALSE}
source(system.file("examples", "pbmc_multisample_demo.R", package = "LogicComm"))

demo <- run_pbmc_multisample_demo(
  seed = 20260526,
  perturb_lambda = 50,
  lcs_threshold = 0.2,
  rank_threshold_grid = c(0.4, 0.5, 0.6, 0.7, 0.8),
  run_rank_evidence = TRUE,
  verbose = TRUE
)
```

The script uses the current public API for the main workflow:

```{r main_api, eval=FALSE}
result <- LogicComm::logic_run(
  pbmc_list,
  group_info = group_info,
  lr_db = LogicComm::lr_pairs_human,
  rank_threshold = 0.5,
  lcs_threshold = lcs_threshold,
  case_label = "Case",
  ctrl_label = "Ctrl",
  chunk_size = 2000,
  mc_cores = 1,
  min_samples_per_group = 2,
  verbose = verbose
)

comparison <- LogicComm::logic_compare_groups(result)
```

The optional rank-aware evidence layer still uses the rank-specific compatibility
function because it exposes threshold-grid diagnostics directly:

```{r rank_api, eval=FALSE}
reo <- LogicComm::logic_prepare(
  pbmc_list[[sname]],
  genes = LogicComm::logic_get_lr_genes(LogicComm::lr_pairs_human),
  lr_db = NULL,
  rank_threshold = 0.5,
  return_rank = TRUE,
  rank_output = "percentile",
  chunk_size = 2000,
  verbose = FALSE
)

rank_res <- LogicComm::IdentifyRankLogicConsensus(
  reo,
  lr_db = LogicComm::lr_pairs_human,
  threshold_grid = rank_threshold_grid,
  lcs_threshold = 0.01,
  verbose = FALSE
)
```

---

## Inspect recovered target rows

```{r target_rows, eval=FALSE}
demo$target_rows[, intersect(c(
  "lr_pair", "pathway", "case_freq", "ctrl_freq", "asymmetry",
  "case_mean_lcs", "ctrl_mean_lcs", "log2fc_lcs", "fdr_fisher",
  "n_case_avail", "n_ctrl_avail"
), names(demo$target_rows)), drop = FALSE]
```

Expected targeted axes include MIF, GALECTIN, and MHC-I examples:

| Target LR pair | Pathway |
|---|---|
| `MIF_CD74` | MIF |
| `MIF_CD44` | MIF |
| `LGALS3_CD44` | GALECTIN |
| `HLA-A_CD8A` | MHC-I |
| `HLA-B_CD8A` | MHC-I |

A successful run should recover these as Case-enriched under the selected
threshold. The exact numeric values can vary with package version, SeuratData
object version, and random seed, so this document describes the expected direction
instead of hard-coding a release-specific result table.

---

## Rank-aware companion evidence

The optional rank-aware table asks whether targeted genes are placed higher in the
within-cell rank order in Case samples.

```{r rank_rows, eval=FALSE}
demo$rank_target_rows[, intersect(c(
  "lr_pair", "pathway", "case_mean_rank_score", "ctrl_mean_rank_score",
  "delta_rank_score", "log2fc_rank_score",
  "case_mean_binary_lcs", "ctrl_mean_binary_lcs"
), names(demo$rank_target_rows)), drop = FALSE]
```

Interpret this table as an effect-size companion to the binary LCS comparison. It
is not a replacement for biological replication or external validation.

---

## Plotting the comparison

```{r plot_demo, eval=FALSE, fig.width=11, fig.height=5}
comparison <- demo$result$comparison
group_info <- demo$group_info

plot_lcs_bubble(
  comparison,
  top_n = 15,
  min_asymmetry = 0.2,
  color_by = "asymmetry"
)

plot_lcs_heatmap(
  comparison,
  group_info = group_info,
  top_n = 20,
  fdr_cutoff = 1
)
```

With only two pseudo-samples per group, FDR will usually be uninformative. In this
demo, focus on `case_freq`, `ctrl_freq`, `asymmetry`, and `log2fc_lcs`.

---

## How this aligns with the core package

- The main workflow uses `logic_run()` rather than the legacy direct
  `run_multisample()` entry point.
- Sample-level comparison can be retrieved with `logic_compare_groups()`.
- REO construction in the optional rank layer uses `logic_prepare()` rather than
  calling `calc_REO_matrix()` directly.
- No Seurat KNN/SNN graph is used for the main ordinary scRNA-seq scoring path;
  LogicComm 0.12 uses sample-level and cell-type co-expression logic.
- The statistical unit is the biological sample. Cells are never treated as
  independent Case/Ctrl replicates.

**中文说明。** 新文档和脚本都改成以 `logic_*` 为入口。KNN/SNN 不再作为普通 PBMC
scRNA-seq 的主证据；多样本比较仍然以 sample 为单位。

---

## Real cohort recommendations

For a real PBMC cohort:

1. Use one object or matrix per biological sample.
2. Keep sample names unique and aligned to `group_info`.
3. Use the same LR database, `rank_threshold`, assay/layer, and preprocessing
   choices across samples.
4. Prefer `logic_run()` for sample-level LR comparison.
5. For cell-type-resolved differential communication, build `ct_comm` per sample,
   convert with `celltype_comm_to_lcs()`, then compare with `logic_compare_groups()`.
6. Interpret p-values only when group sample sizes are sufficient.
7. Support mechanistic claims with specificity, permutation/null diagnostics,
   receiver-response evidence, and independent validation.

---

## Session info

```{r sessioninfo, eval=FALSE}
sessionInfo()
```
