---
title: "Seurat demo: LogicComm 0.12 cell-type co-expression workflow"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  message = FALSE,
  warning = FALSE
)
```

## Overview / 概览

This tutorial shows the recommended **LogicComm 0.12** workflow on a Seurat
PBMC3K object. The main public API is the `logic_*` pipeline:

1. `logic_import_lrdb()` / `logic_check_lrdb()` validate the ligand-receptor database.
2. `logic_prepare()` converts expression to REO binary logic states.
3. `logic_score_lr()` scores sample-level ligand-receptor Logic Consensus Scores (LCS).
4. `logic_summarize_celltypes()` summarizes directed sender-cell-type to
   receiver-cell-type communication.
5. `logic_compare_groups()` or `logic_run()` compares biological samples.

**中文说明。** 本教程按 LogicComm 0.12 的主 API 编写。旧函数
`calc_REO_matrix()`、`IdentifyLogicConsensus()`、`CompareLogicGroups()` 仍保留给历史脚本，
但新分析建议从 `logic_*` 流程开始。

### What changed from older tutorials?

LogicComm now treats dissociated scRNA-seq communication as a **cell-type
co-expression** problem. Seurat's PCA/UMAP/clustering remain useful for annotation
and visualization, but Seurat KNN/SNN graphs are no longer the core evidence for
ordinary scRNA-seq communication scoring. A transcriptomic nearest-neighbor graph
is not physical contact evidence, so the package reports cell-type-level LCS,
activity, specificity, role scores, null diagnostics, and optional validation
adapters.

**中文说明。** 对普通单细胞转录组，LogicComm 不再把 Seurat KNN/SNN 当作局部空间邻域证据。
核心单位是 cell-type co-expression：sender 细胞类型中 ligand 活跃、receiver 细胞类型中
receptor 活跃的比例。KNN/SNN 可以用于 Seurat 聚类和 UMAP，但不要把它解释成真实空间接触。

### Key terms / 关键术语

| Term | Meaning |
|---|---|
| REO | Relative Expression Ordering: per-cell binary logic state based on within-cell gene rank. |
| LCS | Logic Consensus Score: ligand/receptor logic support for a sample or cell-type pair. |
| Cell-type co-expression | Sender-type ligand activity x receiver-type receptor activity across cell-type cells. |
| Same-cell-type signaling | `T cell -> T cell` style homotypic/autocrine-like communication potential. |
| Validation adapters | Optional receiver-response, target, TF-switch, or external evidence layers. |

---

## Installation

```{r install, eval=FALSE}
# Core package dependencies
install.packages(c(
  "Matrix", "dplyr", "ggplot2", "ggrepel", "scales", "Rcpp"
))

# Optional packages used in this tutorial
install.packages(c("Seurat", "SeuratObject", "pheatmap", "patchwork"))

# Install a local source tarball; adjust the path if needed.
install.packages("LogicComm_0.12.3.tar.gz", repos = NULL, type = "source")
```

PBMC3K can be installed via SeuratData:

```{r install_pbmc3k, eval=FALSE}
install.packages("remotes")
remotes::install_github("satijalab/seurat-data")
SeuratData::InstallData("pbmc3k")
```

If SeuratData is unavailable, use the manual 10X download section near the end.

---

## Load packages

```{r load_packages, eval=FALSE}
library(LogicComm)
library(Seurat)
library(SeuratData)
library(Matrix)
library(ggplot2)
```

---

## Step 1: Load PBMC3K

```{r load_pbmc3k, eval=FALSE}
SeuratData::data("pbmc3k")
pbmc <- pbmc3k
pbmc
```

PBMC3K is a single healthy peripheral-blood sample. It is useful for demonstrating
within-sample communication discovery, but it is not a Case-vs-Control cohort.

**中文说明。** PBMC3K 是单个健康 PBMC 样本。它适合演示“哪些细胞类型可能发送/接收信号”，
但不能解读成疾病差异通讯。差异分析需要多个独立生物样本。

---

## Step 2: Standard Seurat preprocessing for annotation and visualization

LogicComm's REO/LCS scoring does not require Seurat normalization. We still run a
standard Seurat workflow because clustering and UMAP help interpret the results.

```{r seurat_preprocess, eval=FALSE}
pbmc <- NormalizeData(pbmc, verbose = FALSE)
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000,
                             verbose = FALSE)
pbmc <- ScaleData(pbmc, verbose = FALSE)
pbmc <- RunPCA(pbmc, npcs = 30, verbose = FALSE)
pbmc <- FindNeighbors(pbmc, dims = 1:20, verbose = FALSE)
pbmc <- FindClusters(pbmc, resolution = 0.5, verbose = FALSE)
pbmc <- RunUMAP(pbmc, dims = 1:20, verbose = FALSE)
```

The neighbor graph created above is for Seurat clustering/UMAP context. The
LogicComm core analysis below uses cell-type co-expression rather than the Seurat
KNN/SNN graph.

---

## Step 3: Choose cell labels

For a quick demo, use `seurat_clusters`. For a real analysis, replace this with a
curated annotation column after marker-gene review.

```{r label_column, eval=FALSE}
label_col <- if ("seurat_annotations" %in% colnames(pbmc@meta.data)) {
  "seurat_annotations"
} else {
  "seurat_clusters"
}

celltype_labels <- setNames(
  as.character(pbmc[[label_col]][, 1]),
  colnames(pbmc)
)

table(celltype_labels)
```

**中文说明。** sender/receiver 的解释完全依赖 cell type label。正式分析应使用人工校正后的
细胞类型注释，而不是未经检查的 cluster 编号。

---

## Step 4: Prepare the LR database

The built-in `lr_pairs_human` is a compact human LR seed database. Validate it once
before analysis:

```{r lrdb_check, eval=FALSE}
data(lr_pairs_human)
logic_check_lrdb(lr_pairs_human, require_metadata = TRUE)
lr_genes <- logic_get_lr_genes(lr_pairs_human)
head(lr_genes)
```

To use CellChatDB, convert it into LogicComm's explicit list-column schema:

```{r cellchatdb_convert, eval=FALSE}
data(CellChatDB.human, package = "CellChat")
cellchat_lr <- logic_import_lrdb(
  CellChatDB.human,
  source = "cellchat",
  include_modulators = TRUE
)
logic_check_lrdb(cellchat_lr, require_metadata = TRUE)
```

Use one LR database consistently across all samples if you plan to compare groups.

---

## Step 5: Build REO logic with `logic_prepare()`

REO converts each cell into binary gene-activity states. `gene_background =
"quantile"` adds a cell-type-free ambient guard and is the recommended default for
PBMC-style data.

```{r logic_prepare, eval=FALSE}
reo_pbmc <- logic_prepare(
  pbmc,
  lr_db          = lr_pairs_human,
  rank_threshold = 0.5,
  layer          = "counts",
  gene_background = "quantile",
  return_rank    = TRUE,
  verbose        = TRUE
)

reo_logic <- reo_pbmc$logic
reo_logic
```

Important points:

- Input matrices are interpreted as **genes x cells**.
- `logic_prepare()` keeps LR genes for output/scoring, while the REO anchor can use
  the full expression background.
- `rank_threshold = 0.5` uses the within-cell median as the activity anchor.
- `return_rank = TRUE` keeps rank-percentile information for rank-weighted
  cell-type scoring.

---

## Step 6: Score sample-level LR logic with `logic_score_lr()`

`logic_score_lr()` returns a named LCS vector for the whole sample. It is useful for
quick LR prioritization and multi-sample comparison.

```{r logic_score_lr, eval=FALSE}
lcs_pbmc <- logic_score_lr(
  reo_pbmc,
  lr_db         = lr_pairs_human,
  lcs_threshold = 0.01,
  verbose       = TRUE
)

head(sort(lcs_pbmc[!is.na(lcs_pbmc)], decreasing = TRUE), 20)
```

Convert the vector to a table when you want to export or merge it:

```{r lcs_table, eval=FALSE}
lcs_tbl <- data.frame(
  lr_pair = names(lcs_pbmc),
  lcs     = as.numeric(lcs_pbmc),
  stringsAsFactors = FALSE
)
lcs_tbl <- lcs_tbl[!is.na(lcs_tbl$lcs), ]
lcs_tbl <- lcs_tbl[order(lcs_tbl$lcs, decreasing = TRUE), ]
head(lcs_tbl, 20)
```

**中文说明。** sample-level LCS 是全样本 LR 逻辑优先级列表，不包含 sender/receiver
细胞类型。要解释“谁发给谁”，继续做 cell-type summary。

---

## Step 7: Score per-cell communication activity

`score_lr_activity()` estimates how many LR programs each cell can participate in
as sender and/or receiver.

```{r score_cells, eval=FALSE}
cell_scores <- score_lr_activity(
  reo_logic,
  lr_db     = lr_pairs_human,
  mode      = "both",
  aggregate = TRUE,
  verbose   = TRUE
)

summary(cell_scores$comm_score)
head(rank_comm_cells(cell_scores, n = 20), 20)
```

Add the scores back to Seurat metadata by cell name:

```{r add_metadata, eval=FALSE}
cell_order <- colnames(pbmc)
pbmc$LogicComm_sender   <- cell_scores$sender_score[cell_order]
pbmc$LogicComm_receiver <- cell_scores$receiver_score[cell_order]
pbmc$LogicComm_score    <- cell_scores$comm_score[cell_order]
```

Visualize on UMAP:

```{r umap_logiccomm, eval=FALSE, fig.width=10, fig.height=4}
FeaturePlot(
  pbmc,
  features  = c("LogicComm_sender", "LogicComm_receiver", "LogicComm_score"),
  reduction = "umap",
  ncol      = 3
)

VlnPlot(pbmc, features = "LogicComm_score", group.by = label_col, pt.size = 0)
```

Interpret high per-cell scores as broad LR logic participation, not as proof that a
cell is biologically more important in every sense.

---

## Step 8: Summarize sender-cell-type -> receiver-cell-type communication

`logic_summarize_celltypes()` is the main bridge from LR scores to biological
interpretation. It reports cell-type-pair summaries, LR drivers, pathway summaries,
role scores, confidence labels, and evidence support.

```{r summarize_celltypes, eval=FALSE}
ct_comm <- logic_summarize_celltypes(
  reo_pbmc,
  seurat_obj      = pbmc,
  label_col       = label_col,
  lr_db           = lr_pairs_human,
  include_self    = TRUE,
  lcs_threshold   = 0.01,
  min_edges       = 20,
  min_active_edges = 1,
  min_expr_frac   = 0.1,
  lcs_weighting   = "rank",
  verbose         = TRUE
)

ct_comm
head(ct_comm$pair_summary[order(-ct_comm$pair_summary$sum_lcs), ], 20)
head(ct_comm$lr_table[order(-ct_comm$lr_table$lcs), ], 20)
```

How to read the key outputs:

| Object | Use |
|---|---|
| `ct_comm$lr_table` | One row per sender type -> receiver type -> LR pair. |
| `ct_comm$pair_summary` | Aggregated sender -> receiver strength and active LR counts. |
| `ct_comm$pathway_summary` | Pathway-level sender -> receiver summaries. |
| `ct_comm$role_summary` | Sender, receiver, mediator, and influencer tendencies per cell type. |

`include_self = TRUE` keeps same-cell-type communication such as `T cell -> T
cell`, interpreted as homotypic or autocrine-like potential. It does not prove that
one single cell both secreted and received the signal.

---

## Step 9: Cell-type communication plots

```{r celltype_plots, eval=FALSE, fig.width=9, fig.height=6}
plot_celltype_heatmap(ct_comm, metric = "sum_lcs")
plot_celltype_heatmap(ct_comm, metric = "n_active_lr")

plot_celltype_roles(ct_comm)
plot_celltype_role_heatmap(ct_comm)

plot_lr_bubble_advanced(ct_comm, top_n_per_pair = 5)
plot_pathway_heatmap(ct_comm, metric = "sum_lcs", top_n_pathways = 25)
```

For one event, inspect the evidence behind the sender -> receiver -> LR claim:

```{r explain_event, eval=FALSE, fig.width=7, fig.height=5}
# Replace these with values present in ct_comm$lr_table.
ex <- explain_celltype_interaction(
  ct_comm,
  sender     = ct_comm$lr_table$sender_type[1],
  receiver   = ct_comm$lr_table$receiver_type[1],
  lr_pair    = ct_comm$lr_table$lr_pair[1],
  reo_mat    = reo_logic,
  seurat_obj = pbmc
)

ex$evidence
ex$plot
```

---

## Step 10: Specificity, ranking, and discovery view

Raw LCS can be dominated by broad immune-recognition axes. Use specificity and the
discovery view to prioritize pair-specific mechanisms.

```{r discovery_ranking, eval=FALSE}
ct_comm <- score_communication_specificity(ct_comm)

ranked_axes <- rank_communication_axes(ct_comm)
head(ranked_axes[, c(
  "sender_type", "receiver_type", "lr_pair", "lcs",
  "specificity_class", "discovery_score", "evidence_tier",
  "broad_axis_flag"
)], 20)

plot_communication_discovery(ranked_axes)

clean_axes <- communication_discovery_view(ranked_axes)
head(clean_axes[, c(
  "sender_type", "receiver_type", "lr_pair", "lcs",
  "discovery_score", "specificity_class", "evidence_tier"
)], 20)
```

**中文说明。** 不建议只按原始 `lcs` 排名。MHC/MIF/B2M 等广谱轴可能很强但不一定是特异机制。
`communication_discovery_view()` 会给出更适合发表机制筛选的视图。

---

## Step 11: Null diagnostics and threshold sensitivity

Permutation tests shuffle cell-type labels to ask whether a sender -> receiver
structure is stronger than random labels. Use enough permutations for the p-value
resolution you need.

```{r permutation_null, eval=FALSE}
null_pair <- permute_celltype_communication(
  ct_comm = ct_comm,
  reo_mat = reo_logic,
  n_perm  = 1000,
  n_cores = 8,
  metric  = "sum_lcs",
  seed    = 1
)

head(null_pair[order(null_pair$fdr), ], 10)
diagnose_permutation_resolution(null_pair)
```

Threshold sensitivity checks whether candidate axes are robust to the REO anchor:

```{r threshold_sensitivity, eval=FALSE}
sens <- sensitivity_REO_threshold(
  pbmc,
  rank_threshold_grid = c(0.4, 0.5, 0.6, 0.7),
  lr_db       = lr_pairs_human,
  cell_labels = celltype_labels,
  layer       = "counts"
)

plot_specificity_stability(sens, ct_comm = ct_comm)
```

With `n_perm = 12`, the smallest empirical p-value is `1 / 13 = 0.077`; use
`n_perm >= 200` or adaptive refinement for interpretable FDR.

---

## Step 12: Optional receiver-response validation

L-R co-expression is communication potential. Receiver-response scoring asks
whether receiver cells also show downstream response-state logic.

```{r receiver_response, eval=FALSE}
response_db <- data.frame(
  lr_pair = c("IFNG_IFNGR1_IFNGR2", "TGFB1_TGFBR1_TGFBR2"),
  response_genes = c("STAT1;IRF1;CXCL10", "SMAD7;SERPINE1;COL1A1"),
  stringsAsFactors = FALSE
)

ct_comm <- add_receiver_response_score(
  ct_comm,
  reo_mat       = reo_logic,
  response_db   = response_db,
  response_mode = "any"
)

head(ct_comm$lr_table[order(ct_comm$lr_table$response_integrated_score,
                            decreasing = TRUE), ], 10)
```

Validation columns are evidence annotations. Count columns alone are not treated as
positive validation support; use positive scores or supportive labels.

---

## Step 13: Multi-sample comparison

For Case-vs-Control studies, one biological sample is the statistical unit. Build
one object per sample, score each sample, then compare sample-level vectors.

```{r multisample_logic_run, eval=FALSE}
pbmc_list <- list(
  sample001 = pbmc_case_1,
  sample002 = pbmc_case_2,
  sample003 = pbmc_ctrl_1,
  sample004 = pbmc_ctrl_2
)

group_info <- c(
  sample001 = "Case",
  sample002 = "Case",
  sample003 = "Ctrl",
  sample004 = "Ctrl"
)

analysis <- logic_run(
  pbmc_list,
  group_info     = group_info,
  lr_db          = lr_pairs_human,
  rank_threshold = 0.5,
  lcs_threshold  = 0.01,
  layer          = "counts",
  chunk_size     = 2000,
  mc_cores       = 1,
  verbose        = TRUE
)

comparison <- logic_compare_groups(analysis)
head(comparison[order(comparison$fdr_fisher), ], 20)
```

For cell-type-resolved cohort comparison, compute `ct_comm` separately for each
sample and compare vectors from `celltype_comm_to_lcs()`:

```{r multisample_celltype, eval=FALSE}
sample_ct_list <- lapply(names(pbmc_list), function(sname) {
  reo_i <- logic_prepare(
    pbmc_list[[sname]],
    lr_db          = lr_pairs_human,
    rank_threshold = 0.5,
    layer          = "counts",
    gene_background = "quantile",
    return_rank    = TRUE,
    verbose        = FALSE
  )

  labels_i <- setNames(
    as.character(pbmc_list[[sname]][[label_col]][, 1]),
    colnames(pbmc_list[[sname]])
  )

  logic_summarize_celltypes(
    reo_i,
    cell_labels     = labels_i,
    lr_db           = lr_pairs_human,
    lcs_threshold   = 0.01,
    min_edges       = 20,
    min_expr_frac   = 0.1,
    lcs_weighting   = "rank",
    verbose         = FALSE
  )
})
names(sample_ct_list) <- names(pbmc_list)

ct_lcs_list <- lapply(sample_ct_list, celltype_comm_to_lcs,
                      level = "celltype_lr", metric = "lcs")

ct_comparison <- logic_compare_groups(
  ct_lcs_list,
  group_info    = group_info[names(ct_lcs_list)],
  case_label    = "Case",
  ctrl_label    = "Ctrl",
  lcs_threshold = 0.01
)

plot_differential_celltype_heatmap(ct_comparison, metric = "asymmetry")
plot_differential_celltype_volcano(ct_comparison, x = "log2fc_lcs")
```

Do not merge all cells from all samples and treat them as replicates. That creates
pseudo-replication.

---

## Step 14: Manual PBMC3K download alternative

```{r manual_download, eval=FALSE}
dir.create("data", showWarnings = FALSE)
url <- "https://cf.10xgenomics.com/samples/cell-exp/1.1.0/pbmc3k/pbmc3k_filtered_gene_bc_matrices.tar.gz"
dest <- file.path("data", "pbmc3k_filtered_gene_bc_matrices.tar.gz")
download.file(url, destfile = dest, mode = "wb")
untar(dest, exdir = "data")

pbmc_data <- Read10X(data.dir = "data/filtered_gene_bc_matrices/hg19")
pbmc <- CreateSeuratObject(counts = pbmc_data, project = "pbmc3k")
```

Then continue from Step 2.

---

## Troubleshooting

### `logic_summarize_celltypes()` reports missing labels

Make sure labels are named by cell barcode and cover all REO cells:

```{r troubleshoot_labels, eval=FALSE}
all(colnames(reo_logic) %in% names(celltype_labels))
setdiff(colnames(reo_logic), names(celltype_labels))[1:5]
```

### Memory usage is high

Lower `chunk_size` in `logic_prepare()`:

```{r troubleshoot_memory, eval=FALSE}
reo_pbmc <- logic_prepare(pbmc, lr_db = lr_pairs_human,
                          layer = "counts", chunk_size = 1000)
```

### Which Seurat layer should I use?

Raw counts are a good default for REO logic:

```{r troubleshoot_layer, eval=FALSE}
reo_pbmc <- logic_prepare(pbmc, lr_db = lr_pairs_human, layer = "counts")
```

Keep the same layer across all samples in a cohort.

### Why not use the Seurat KNN graph as core evidence?

For dissociated scRNA-seq, KNN/SNN edges represent transcriptional similarity, not
physical proximity. LogicComm keeps the core inference at the cell-type
co-expression level and reserves spatial/neighbor graphs for appropriate spatial or
legacy compatibility contexts.

---

## Session info

```{r sessioninfo, eval=FALSE}
sessionInfo()
```
